### 2006-02-19 paper/trunk/systemfit.tex
revised abstract and keywords
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@615 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%       $Id$

\documentclass[article]{jss}
\usepackage[pdftex]{thumbpdf}
%   run: pdflatex *.tex -> thumbpdf *.tex -> pdflatex *.tex
\usepackage{csquotes}
\MakeOuterQuote{°}
\usepackage{amsmath}
\usepackage{float}

\usepackage{lineno}

\linenumbers

\newcommand{\Cov}{\text{Cov}}
\clubpenalty=10000
\widowpenalty=10000

\author{Arne Henningsen\\University of Kiel\And
        Jeff D. Hamann\\ Forest Informatics, Inc. }

\title{systemfit: A Package to Estimate\\
       Simultaneous Equation Systems in \proglang{R}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Arne Henningsen, Jeff D.\ Hamann} %% comma-separated
\Plaintitle{systemfit: A Package to Estimate Simultaneous Equation Systems 
            in R} %% without formatting
\Shorttitle{systemfit} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{ 
Many statistical analyses are based on models containing systems of
structurally related equations.
Since in most cases cross-equation disturbances are correlated,
full information methods are required \citep{zellner62}.
If exogenous variables are stochastically dependent
on the disturbances in the system,
instrumental variable estimations should be used for parameter estimation
\citep{zellner62b}
The package \pkg{systemfit} provides the capability
to estimate systems of linear equations
within the \proglang{R} programming environment.
}

\Keywords{R, simultaneous equations systems,
  seemingly unrelated regression, two-stage least squares,
  three-stage least squares}

%% at least one keyword must be supplied

%% publication information
%% NOTE: This needs to filled out ONLY IF THE PAPER WAS ACCEPTED.
%% If it was not (yet) accepted, leave them commented.
%% \Volume{13}
%% \Issue{9}
%% \Month{September}
%% \Year{2004}
%% \Submitdate{2004-09-29}
%% \Acceptdate{2004-09-29}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
   Arne Henningsen\\
   Department of Agricultural Economics\\
   University of Kiel\\
   D-24098 Kiel, Germany\\
   E-mail: \email{ahenningsen@agric-econ.uni-kiel.de}\\
   URL: \url{http://www.uni-kiel.de/agrarpol/ahenningsen/}\\
   \\
   Jeff D.\ Hamann\\
   Forest Informatics, Inc.\\
   PO Box 1421\\
   Corvallis, Oregon 97339-1421\\
   e-mail: \email{jeff.hamann@forestinformatics.com}\\
   URL: \url{http://www.forestinformatics.com}\\
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/1/31336-5053
%% Fax: +43/1/31336-734

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\input{systemfit_intro.tex}
\input{systemfit_statistics.tex}
\input{systemfit_usage.tex}
\input{systemfit_summary.tex}

% \section{CVS Revision Log}
% \begin{verbatim}
% $Log$
% Revision 1.19  2005/11/08 06:31:48  hamannj
% made seemingly massive changes... but hopefully not for the worse...
%
% Revision 1.18  2005/11/07 14:08:56  henningsena
% a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
%
% Revision 1.17  2005/11/07 10:21:59  henningsena
% just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
%
% Revision 1.16  2005/11/02 18:31:45  hamannj
% small last changes
%
% Revision 1.15  2005/11/02 18:15:40  hamannj
% small changes

% Revision 1.14  2005/10/26 21:16:53  hamannj
% checked out and updated local copy before edits

% Revision 1.13  2005/09/26 20:34:12  henningsena
% using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)

% Revision 1.12  2005/09/26 11:24:05  henningsena
% some minor changes

% Revision 1.11  2005/09/01 12:27:55  henningsena
% revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.

% Revision 1.10  2005/09/01 11:49:02  henningsena
% moved section about non-linear estimation from systemfit.tex into new file nlsystemfit.tex, because this section should not be included in (this version of) the article

% Revision 1.9  2004/11/16 10:59:09  henningsena
% some minor changes and some comments on nlsystemfit

% Revision 1.8  2004/11/16 08:55:53  henningsena
% merged file and resolved conflict due to some minor changes I did while Jeff was adding nlsystemfit.

% Revision 1.7  2004/11/16 05:34:47  hamannj
% question about nlsystemfit code section (example)

% Revision 1.6  2004/11/16 05:32:46  hamannj
% made small changes because my code section wouldn't compile. argh.

% Revision 1.5  2004/11/16 05:31:19  hamannj
% added cvs header and log section. to be removed when submitted.

% \end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\bibliography{systemfit} % a subset of my big bibtex file
\bibliography{/home/suapm095/Documents/Literatur/arne}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}