### 2006-02-25 paper/trunk/systemfit.tex
added footer for nested docs
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@632 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%       $Id$

\documentclass[article]{jss}
\usepackage[pdftex]{thumbpdf}
%   run: pdflatex *.tex -> thumbpdf *.tex -> pdflatex *.tex
\usepackage{csquotes}
\MakeOuterQuote{°}
\usepackage{amsmath}
\usepackage{float}

\usepackage{lineno}

\linenumbers

\newcommand{\Cov}{\text{Cov}}
\clubpenalty=10000
\widowpenalty=10000

\hyphenation{sys-tem-fit}

\author{Arne Henningsen\\University of Kiel\And
        Jeff D. Hamann\\ Forest Informatics, Inc. }

\title{systemfit: A Package to Estimate\\
       Simultaneous Equation Systems in \proglang{R}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Arne Henningsen, Jeff D.\ Hamann} %% comma-separated
\Plaintitle{systemfit: A Package to Estimate Simultaneous Equation Systems 
            in R} %% without formatting
\Shorttitle{systemfit} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{ 
Many statistical analyses are based on models containing systems of
structurally related equations.
In cases where cross-equation disturbances are correlated,
full information methods are required \citep{zellner62}.
If exogenous variables are stochastically dependent
on the disturbances in the system,
then instrumental variable estimation methods should be used
\citep{zellner62b}
The package \pkg{systemfit} provides the capability
to estimate systems of linear equations
within the \proglang{R} programming environment.
}

\Keywords{R, simultaneous equations systems,
  seemingly unrelated regression, two-stage least squares,
  three-stage least squares}

%% at least one keyword must be supplied

%% publication information
%% NOTE: This needs to filled out ONLY IF THE PAPER WAS ACCEPTED.
%% If it was not (yet) accepted, leave them commented.
%% \Volume{13}
%% \Issue{9}
%% \Month{September}
%% \Year{2006}
%% \Submitdate{2006-02-28}
%% \Acceptdate{2006-09-29}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
   Arne Henningsen\\
   Department of Agricultural Economics\\
   University of Kiel\\
   D-24098 Kiel, Germany\\
   E-mail: \email{ahenningsen@agric-econ.uni-kiel.de}\\
   URL: \url{http://www.uni-kiel.de/agrarpol/ahenningsen/}\\
   \\
   Jeff D.\ Hamann\\
   Forest Informatics, Inc.\\
   PO Box 1421\\
   Corvallis, Oregon 97339-1421\\
   e-mail: \email{jeff.hamann@forestinformatics.com}\\
   URL: \url{http://www.forestinformatics.com}\\
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/1/31336-5053
%% Fax: +43/1/31336-734

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\input{systemfit_intro.tex}
\input{systemfit_statistics.tex}
\input{systemfit_code.tex}
\input{systemfit_usage.tex}
\input{systemfit_reliability.tex}
\input{systemfit_summary.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{systemfit} % a subset of my big bibtex file
%\bibliography{/home/suapm095/Documents/Literatur/arne}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: