### 2006-02-25 paper/trunk/systemfit_summary.tex
added "\nopagebreak" to avoid a page break directly after the section header
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@637 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%       $Id$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary and outlook}\label{sec:Summmary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\nopagebreak
In this article, we have described some of the basic features of the
\pkg{systemfit} package for estimation of linear systems of
equations with various estimation methods.
Many details of the estimation can be controlled by the user.
Furthermore, the package provides some statistical tests
for parameter restrictions and consistency of 3SLS estimation.
It has been tested on a variety of datasets and has produced satisfactory
for a few years.
While the \pkg{systemfit} package performs the basic fitting methods,
more sophisticated tools exist.
We hope to implement missing functionalities
in the near future.
Some of these are discussed in the following.

\subsubsection*{Unbalanced datasets}
Currently, the \pkg{systemfit} package requires
that all equations have the same number of observations.
However, many data sets have unbalanced observations.%
\footnote{
For instance,
forestry datasets typically contain observations of many inexpensive
observations (stem diameter) and few expensive observations such
as stem height.
}
Simply dropping data points
that do not contain observations for all equations
may considerably reduce the number of observations,
and thus, the information utilized in the estimation.
Hence, it is our intention to include the capability for estimations
with unbalanced data sets as described in \citet{schmidt77}
in future releases of \pkg{systemfit}.

\subsubsection*{Serial correlation and heteroscedasticity}
For all of the methods developed in the package, the disturbances of
the individual equations are assumed to be independent and identically
distributed (iid).
The package could be enhanced by the inclusion of methods to fit
equations with serially correlated and heteroscedastic disturbances.

\subsubsection*{Estimation methods}
We wish to include more sophisticated estimation methods such
as limited-information maximum likelihood (LIML),
full-information maximum likelihood (FIML),
and generalized methods of moments (GMM), in future releases.

\subsubsection*{Non-linear estimation}
The \pkg{systemfit} package provides a function to estimate
systems of non-linear estimations.
However, the function \code{nlsystemfit} is currently under development
and the results are not yet always reliable
due to convergence problems.
We hope to improve this behavior soon.

%% testing for the new svn repos
%% It works great!!!

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "systemfit"
%%% End: