### 2006-08-28 paper/trunk/JSS/response_A.tex
JSS/reviewer_*.*: added reviews
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@682 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\documentclass[12pt,english]{scrartcl}
\input{response_commands}

\begin{document}
\begin{center}
{\Large\textbf{
systemfit: A Package to Estimate
Simultaneous Equation Systems in R (JSS~224)}}

\bigskip

{\Large\textbf{Responses to Reviewer A}}
\end{center}

We thank this reviewer for a very helpful set of comments and suggestions
for improving the paper.
The reviewer has made one overall comment, 9 major comments and
several minor comments.
We have structured our reply accordingly.

\section{Overall Comment}
% ########################################################
% ## Review for                                         ##
% ##   systemfit: A Package to Estimate Simultaneous    ##
% ##   Equation Systems in R                            ##
% ##   Arne Henningsen, Jeff D. Hamann                  ##
% ## submitted for publication in                       ##
% ##   Journal of Statistical Software                  ##
% ########################################################
% 
% The paper describes the CRAN package "systemfit" for fitting simultaneous
% equation models in R. Estimation procedures include standard techniques in
% econometrics such as SUR, 2SLS and 3SLS.
% The paper can be divided roughly into three parts: Sections 1/2 explain
% the underlying theory (~12 pages), Sections 3/4 describe the code in systemfit
% and how to use it (~10 pages) and Section 5 reproduces results from various
% textbooks (~8 pages).
% The paper and software are both interesting and mostly well-written: they
% deserve publication in JSS. However, some major and minor comments have to
% be addressed before it can be accepted. In particular, the revision has to
% streamline exposition and notation of standard results, and to combine the
% (established) theory with the implementation more closely (otherwise the reader
% has to wait more than 10 pages before s/he learns something about the package).
% I suggest to eliminate Section 2 in its current form and to re-write the
% description of the theory in a more general and unified way while weaving
% it with the translation into computational tools. As for the software itself,
% several modifications are necessary to make it more intuitive for R users
% and to improve how the underlying concepts are turned into computational 
% tools. More details are described below.
% 
% 
\section{Major Points}
% MAJOR POINTS:
\subsection{Title}
%   o title
% 
%     The title sounds a bit awkward in English: first, "a package *for*
%     estimat*ing*" would be more appropriate, secondly "systems of
%     equations" seems to be preferred over "equation systems" in English.
% 
\subsection{Abstract/Introduction}
%   o abstract/introduction
% 
%     Both, abstract and introduction, talk rather generally about
%     systems of equations reviewing rather well-known theoretical
%     results. I would suggest to revise both, focusing on what
%     *can* be done *in R* rather than giving a tutorial what *should*
%     be done *in general*. Of course, it is ok to review some theory,
%     however, practitioners should know what they are doing and the
%     paper should mainly teach them how to do it in R.
%
\subsection{Estimation Methods (2.1)}
%   o estimation methods (2.1)
% 
%     The description of the estimation methods can be shortened considerably
%     and should focus on the unifying properties of the different estimators.
%     All methods from OLS through to standard 3SLS comply with formulas
%     (12) and (13), only with differing choices for Omega (s*I, diagonal or
%     full Omega) and X (original X vs \hat X from instruments). The exposition
%     could be streamlined considerably if this structure were exploited.
% 
%     Terminology for the different estimators also varies between various
%     communities using systems of equations models. The paper should address
%     this. For example, in econometrics the GLS estimator would be an estimator
%     with known Omega while FGLS (feasible GLS) is the estimator based on
%     an estimated Omega (see Greene, 2003). The current discussion makes this
%     even more problematic because Omega is treated as known throughout
%     Section 2. The discussion of the estimators needs to be combined with
%     discussion of the estimators for Omega.
% 
%     For the varying 3SLS estimators, it would help to point out why/when
%     these are needed at the beginning rather than at the end of the 3SLS
%     section. 
%
\subsection{Imposing Linear Restrictions (2.2)}
%   o imposing linear restrictions (2.2)
% 
%     I found this section somewhat confusing. At first, it should be made
%     clear that this is about estimating the coefficients under linear
%     restrictions (and not inference or something else). Then, it needs
%     to be explained more concisely that all linear restrictions are of
%     type (33) and that the rest of the discussion is about different
%     computational algorithms for obtaining the estimates under restriction.
%     Again, the exposition on pages 7-9 can be improved (and shortened)
%     if the unifying structure is exploited. Also the notation can be
%     streamlined concerning beta, \hat beta, beta^0, beta^* etc.
%
\subsection{systemfitClassic (4.3)}
%   o systemfitClassic (4.3)
% 
%     First, this function deserves a better name: it is long, unintuitive
%     and does not reflect the underlying conceptual tools. If the requirement
%     of using the same formula for all equations would be relaxed (but
%     still supported) a conceivable name would be sur() for SUR type 
%     models. But it would also be conceivable to stick to the current
%     requirement of a single formula (then, still a better name would
%     be needed) and to set up a second wrapper sur() etc. In any case,
%     the function (its name, arguments and output) should reflect what
%     is conceptually intuitive for this kinds of models (for the communities
%     using it).
%     Second, the interface should also be re-designed: the standard in R is
%     that the formula/s is/are the first argument (which would also be more
%     convenient for systemfit()), e.g.,
%       function(formula, method = "FGLS", data = ..., ...)
%     Currently, the data argument seems to be mandatory (opposed to the
%     paper documentation). The specification of eqnVar and timeVar via
%     strings seems to be uncommon and suboptimal. Probably some kind of
%     formula would be preferable. Maybe a formula similar to those used in
%     nlme could be employed? If more than one formula is specified, one has
%     to pay attention that no problems are used by different patterns of
%     missingness in differing equations.
%
\subsection{Testing Linear Restrictions (4.4)}
%   o testing linear restrictions (4.4)
% 
%     The package introduces several functions that are somewhat orthogonal
%     to exisiting methodology. A single method (with "test" argument) for
%     the linear.hypothesis() function from "car" might seem more appropriate
%     here. The Wald test can already now be reproduced using the default
%     linear.hypothesis() method (except for formula formatting)
%       linear.hypothesis(fitsur, Rmat, qvec)    
%     If the coefficient naming would be modified (see below), the convenience
%     interface for specifying the restrictions in linear.hypothesis() could
%     also be re-used.    
%     Also two of the functions look like methods to the generics lrtest() and
%     waldtest() in package "lmtest", although this is not the case. Furthermore,
%     some more general method of comparing various fitted models would be nice,
%     e.g., an anova() method or methods to waldtest()/lrtest() although the 
%     latter two seem to be under development.
%     The output formatting of the current function is also not very elegant,
%     a new class along with its print method is introduced for each test. It
%     would be much easier to re-use "htest" or (probably better) "anova" as
%     classes for this.
%
\subsection{Organize Internals of systemfit}
%   o organize internals of systemfit
% 
%     The internals of a fitted "systemfit" object could be re-organized. Currently,
%     there are a lot of single letter elements - more standard naming as in "lm"
%     objects etc. would be more desirable. Also, computing the partial t tests
%     should/could be done in the summary() method, avoiding that certain parameters
%     (not affecting the fit itself, such as degrees of freedom) necessarily have to
%     be specified for model fitting.
% 
%     The technical control parameters could also be collected in a list of arguments, e.g.,
%       control = systemfit.control()
%     that sets suitable defaults but can also easily be changed. If possible `...'
%     arguments could be passed to systemfit.control() assuring backward compatibility
%     and facilitating interactive usage. The names of various
%     arguments should be re-thought, e.g. "tx", "R.restr" and "q.rests" seem rather
%     arbitrary. Similar comments hold for other arguments, see also below for some
%     more comments. As noted above, the standard in R seems to supply the formula(s)
%     as the first argument and to have then some `method' argument.
% 
%     Some more standard elements of a model fit should also be kept, in particular
%     the call so that update() can work with "systemfit" objects, e.g.,
%       fitsur <- systemfit("SUR", list(demand = eqDemand, supply = eqSupply))
%       fitsur$call <- call("systemfit", "SUR", list(demand = eqDemand, supply = eqSupply))
%       fitsurRmat <- update(fm, R.restr = Rmat, q.restr = qvec)
%     The call can easily be captured via match.call(). Similarly, one could think 
%     about how the terms and formula can be preserved and possibly made available
%     in terms() and formula() calls. Other standard methods would also be nice to have,
%     e.g., df.residual() or logLik(). (By the way: The row/column naming in fitted()
%     and residuals() should match, probably re-using equation and observation names.)
% 
%     The coefficient naming should be improved, e.g., "<eqname>_<varname>"
%     where <eqname> can be set by default to "eq1", "eq2" etc. but should use the
%     equation names (if supplied). Spaces in names should be avoided.
%     The coefficient naming in the systemfitClassic() interface could also be 
%     improved, see e.g. the output on p.27, l.546: "eq 2 capital.Chrysler" redundantly
%     codes the firm (both by eq and name) and "Chrysler_capital" or "Chryler.capital"
%     would be preferable.
% 
%     It would also be nice to have some more control about the amount of output of
%     the summary() method. Currently, the output can become very long easily when the
%     number of equations increases. One could want to control the equations and
%     correlations used, e.g., by something like
%       summary(sfobject, equation = ..., cor = ...)
% 
%
\subsection{Computations in systemfit}
%   o computations in systemfit()
% 
%     A few computations within the systemfit() functions can be further streamlined.
%     A simple modification is to use
%       method %in% c("OLS", "WLS", ...)
%     instead of
%       method == "OLS" | method == "WLS" | method == ...
%     Furthermore, various marix computations can be made even more stable. Currently,
%     the code still has    
%       solve(crossprod(X), crossprod(X,Y))    
%       solve(crossprod(X))
%     which looks very similar to the code presented in Bates (2004). However, Bates
%     (2004) uses the "Matrix" package, which has specialized solve() for certain
%     matrix types. To perform this `by hand' without Matrix, one way would be re-using
%     lm.fit() for the first command. Alternatively, the QR or Cholesky decomposition
%     could be carried out using qr() or chol(). The latter command could be replaced
%     by
%      chol2inv(qr.R(qr(X)))
%      chol2inv(chol(crossprod(X)))
%     where chol() should be faster but qr() might be more accurate.
%
\subsection{Replication of Textbook Results (5)}
%   o Replication of textbook results (5)
% 
%     Having a section like this is excellent and should help teachers and students
%     for using systemfit in their lectures. Furthermore, it encourages reproducible
%     research which always deserves support. However, I would suggest to streamline
%     the section to make it more readable. The section should discuss its aims (where
%     I would add replication to reliability) and summarize the most important results
%     and experiences from the replication. Then, I would pick a `typical' example and
%     discuss it in some more detail with full R output. For the remaining examples,
%     the input suffices as long as there are no deviations of the systemfit results
%     from the published numbers.
% 
%     In addition to Table 14.1 and 14.2, I tried to replicate Table 14.3 with
%     ML estimates (using iterated FGLS) in Greene (2003) via 
%       systemfitClassic("WSUR", invest ~ value + capital, "firm", "year",
%                        data = GrunfeldGreene, rcovformula = 0, pooled = TRUE, maxiter = 1000)    
%     for the pooled model which yielded rather similar results; and via
%       systemfitClassic("SUR", invest ~ value + capital, "firm", "year",
%                        data = GrunfeldGreene, rcovformula = 0, maxiter = 1000)    
%     for the full model which is not unsimilar to Greene's results but also not very
%     close either. Is this a problem of systemfit or of Greene or of me fitting the
%     wrong model?
% 
%     Having two Grunfeld data sets is confusing as the data used by Theil is simply
%     a subset of the data used by Greene:
%       subset(GrunfeldGreene, firm %in% c("General Electric", "Westinghouse"))
%     Of course, there are different versions of this data set, but neither the paper
%     nor the manual pages of GrunfeldGreene and GrunfeldTheil appropriately discuss
%     the problem. From the manual and the paper, the reader is only able to feel
%     vaguely insecure about the appropriateness of the selected subsets. Also note
%     that a more complete (and slightly different) version of the data set is
%     available in package "Ecdat".
% 
%
\section{Minor Points and Typos}
% MINOR POINTS & TYPOS
% 
%   o p.1, l.3. Is a "theoretical model that is econometrically estimated"
%     a well-defined term?
% 
%   o p.2, Eq.2. The \cdot between the regressor matrix and its coefficients
%     should be omitted (since this is not a scalar multiplication).
% 
%   o p.2, l.42. For symmetry between the two index types, it might be
%     more convenient to denote the equation by i/j and the observation
%     by s/t rather than t/t*, in particular as * is also used later for 
%     other purposes.
% 
%   o p.3, l.45. While it is possible to call u "residual" and \hat u an 
%     "estimated residual", the more standard (and less confusing) terminology
%     is to call u "disturbance" or "error term" and the estimated quantity
%     \hat u "residual".
% 
%   o p.3, l.52. The equation $\sigma^2 = E(u'u)$ is wrong because the 
%     right-hand side is a sum over squared errors. This mistake is repeated
%     throughout the paper. Also (as noted above), no distinction between
%     population and sample covariances is made which leads to confusion.
% 
%   o p.3, l.69. "the GLS is" -> "the GLS estimator is"
%                "called ... SUR" -> "called ... SUR estimator"
% 
%   o p.4, l.72/3. The "E[u|X] = 0" type equations imply (but are not
%     equivalent to) the assumptions of uncorrelated disturbances.
% 
% 
%   o p.4, l.79. The formulation "estimated coefficients are biased" is somewhat
%     sloppy, bias is a property of the estimator.
% 
%   o p.6, l.127 / p.10, l.198. The symbol T is used for different quantities
%     here.
% 
%   o p.12, eq.63. One Sigma is lacking a hat. Now M is used for the number
%     of equations, before it was G.
% 
%   o p.13, l.285. Rather than calling the methods by their full name, it
%     is preferred to refer to the predict or coef method for "systemfit"
%     objects.
% 
%   o p. 13, l.289. The remarks about why not the full code is presented
%     sound very clumsy and should be avoided.
% 
%   o p.13, l.299. "interest" -> "interested"
% 
%   o p.13, l.321. "deviding" -> "dividing"
% 
%   o p.14, l.334. values to arguments should also be set in verbatim,
%     e.g., "``OLS''" -> "\code{"OLS"}"
% 
%   o p.14, l.340. "tilde" -> "\verb/~/"
% 
%   o p.14, l.341. Are only "+" signs allowed or the full standard
%     formula language in R?
% 
%   o p.14, eq.68/69/71. "*" -> "\cdot" (and not in verbatim)
% 
%   o p.14, l.346. "dependant" -> "dependent"
% 
%   o p.17, l.387. The creation of tx should be simplified, e.g., the
%     first six lines could be replaced by
%       tx <- cbind(rbind(diag(5), 0, 0), 0)
% 
%   o p.17, l.405. "large" -> "larger", also say that maxiter = 1 is the
%     default.
% 
%   o p.18, Tab.1. I find it a bit confusing that the argument ist called 
%     "rcovformula" but does not expect a formula (in the R sense). Also,
%     type 0 should have an alternative string specification. A similar
%     argument holds for "formula3sls".
%

% \bibliographystyle{}
% \bibliography{}

\end{document}