### 2006-08-29 paper/trunk/systemfit_statistics.tex
corrected / harmonized punctuation marks around equations (first small comment of reviewer B)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@686 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Statistical background}\label{sec:statistics}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section we provide the statistical background
of the functionality provided by the \pkg{systemfit} package.
After introducing notations and assumptions,
we provide the formulas to estimate systems of linear equations.
We then demonstrate how to impose linear restrictions on parameters.
Finally, we present additional relevant issues about estimation of
equation systems.

Consider a system of $G$ equations, where the $i$th equation is of
the form 
\begin{equation}
   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G ,
\end{equation}
where $y_i$ is a vector of the dependent variable,
$X_i$ is a matrix of the exogenous variables,
$\beta_i$ is the coefficient vector and
$u_i$ is a vector of the disturbance terms of the $i$th equation.

We can write the °stacked° system as
\begin{equation}
   \left[ \begin{array}{c}
      y_1 \\ y_2\\ \vdots\\ y_G
   \end{array} \right] = 
   \left[ \begin{array}{cccc}
      X_1 & 0 & \cdots & 0\\
      0 & X_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & X_G
   \end{array}\right] \cdot 
   \left[ \begin{array}{c}
      \beta_1 \\ \beta_2 \\ \vdots\\ \beta_G
   \end{array} \right] +
   \left[ \begin{array}{c}
      u_1 \\ u_2 \\ \vdots\\ u_G
   \end{array} \right]
\end{equation}

or more simply as
\begin{equation}
   y = X \beta + u .
\end{equation}

We assume that there is no correlation of the disturbance terms 
across observations, so that
\begin{equation}
   E \left( u_{it} \, u_{jt^*} \right) = 0 
   \; \forall \; t \neq t^* ,
\end{equation}
where $i$ and $j$ indicate the equation number 
and $t$ and $t^*$ denote the observation number.

However, we explicitly allow for contemporaneous correlation, i.e.\
\begin{equation}
   E \left( u_{it} \, u_{jt} \right) = \sigma_{ij} .
\end{equation}

Thus, the covariance matrix of the total system is
\begin{equation}
   E \left( u \, u' \right) = \Omega = \Sigma \otimes I ,
\end{equation}
where $\Sigma = \left[ \sigma_{ij} \right]$ is the residual covariance matrix
and $I$ is an identity matrix.

\input{systemfit_estimation.tex}
\input{systemfit_restrictions.tex}
\input{systemfit_other.tex}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "systemfit"
%%% End: