### 2006-10-02 paper/trunk/JSS/response_B.tex
started responding to reviewer B comments. have to go to class for now.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@703 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\documentclass[12pt,english]{scrartcl}
\input{response_commands}

\begin{document}
\begin{center}
{\Large\textbf{
systemfit: A Package to Estimate
Simultaneous Equation Systems in R (JSS~224)}}

\bigskip

{\Large\textbf{Responses to Reviewer A}}
\end{center}

We thank this reviewer for a very helpful set of comments and suggestions
for improving the paper.
The reviewer has made one overall comment, 9 specific comments and
several smaller comments.
We have structured our reply accordingly.

\section{Overall Comment}
% Review of the manuscript 'systemfit: A package to estimate
% simultaneous equation systems in R'
% This manuscript describes a new R package systemfit for inference in multi-equation models. The
% main parts are section 2 and 4. Section 2 gives the statistical background for the analysis tools
% available in systemfit while section 4 describes how to apply systemfit. The theory is adequately
% explained and the syntax in systemfit is nice and logical allowing complicated models to be quite
% easily fitted.
% My only real concern is that systemfit relies on classic econometric (least squares) developed 40
% years ago. Since then computational power has increased dramatically and superior methods may be
% available. First of all, full likelihood methods are available in existing software for structural
% equation modelling. In particular, R has a package called SEM. I am not sure what is gained with
% systemfit compared to this. Also it is a weakness that systemfit only allows analysis of complete
% cases.

\textbf{[Still working on this section]}

It was our intention to create a package for fitting models that could
be verified against published results before moving on to include more
sophisticated fitting methods.

The reasons for doing this were two-fold.
Our first reasons was to simply introduce a econometrics package where
the main focus of the package was econometrics (a branch of the R
packages that was under-represented).

that at the time of the package introduction (2002) was

The second reason for creating the package was to create a "forum" for
more advanced fitting methods as described in Section 6 (Summary and
Outlook). It is our intention to impliment these more advanced methods
over time, but we felt that having a "basic" package that was accepted
as part of a peer-review process would create an environment that
would allow for acceptance in the future since the more sophisticated
methods can be subject to difficulties beyond numerical routines.

\section{Specific comments}
% Specific comments

\subsection{Instrumental variable estimation}
% Instrumental variable estimation has received much attention lately
% in biostatistics. Maybe it would help 'sell' systemfit if this was
% emphasized at bit more. For instance, instrumental variables could
% be mentioned directly in the introduction and included as a key
% word.

We have mentioned "instrumental variables" in the abstract, include
three examples of topics that use IV methods and have added
"instrumental variables" to the keywords.

\subsection{Section 2}
% Section 2 gives a brief overview of the statistical background. This
% may be a bit hard to follow for people outside econometrics. It
% would help to include more references, especially in section 2.1.

\subsection{Number of observations}
% Page 2: 'observations' should be explained earlier together with the presentation of the data
% structure. The number of observations should be stated.

\subsection{x and y variables}
% Page 2: It should be clarified whether y-variables in one equation can be x-variables in another.

\subsection{'may' should be 'must'}
% Page 4, line 82: 'may' should be 'must'. Here the difference is extremely important. Maybe the text
% could emphasize more clearly what is required of a variable if it is to be used as an instrumental
% variable. What is a good instrumental variable and what happens if the requirements are not
% fulfilled?

\subsection{Equation 20 and 23}
% The difference between equation 20 and 23 is so smaller that maybe it should be noted in the text
% what it is.

\subsection{Sigma}
% Line 193-194: Language is confusing. It should be explained that calculation of some of the
% estimators earlier introduced require Sigma to be known. Often Sigma is not known so an estimator
% is used instead. However, then statistical properties may changed compared to the theory described
% earlier.

\subsection{First step OLS}
% Line 196: It should be explained what is meant by a first step OLS.

\subsection{Efficiency and maximum likelihood}
% Section 2.3: Somewhere in this section or (2.1) it should be noted that the efficiency results of
% section 2.1 depend on Sigma being known. And that a full maximum likelihood analysis may be
% superior.
%

\subsection{Alternative approach}
% Line 215-219: Alternative approach should be explained more clearly.

\subsection{Syntax for instrumental variables}
% Line 375-379: Syntax on programming of instrumental variables should be made more clear. What
% is a one-sided formula? Does the code given imply that the sum of income, farmPrice and trend was
% used as and instrumental variable? If so, it should be stated in the text.

\subsection{Section 2.7}
% Section 2.7: Given the theory of the previous sections, I was confused about the Hausman test.
% Under the alternative hypothesis, independent variables are correlated with error terms. In this
% section it is stated that this leads to inconsistency of 3SLS. However, on page 5 (top) the opposite is
% stated? Would it not make more sense to compare SUR to 3SLS?

\section{Smaller comments}
% Smaller corrections:
% Sometimes ':' is used just before an equation other times not.
% Page 7, line 147: change 'doesn't' to 'does not'.
% Line 135: '\beta^*' should be '\widehat{\beta^*}'
% Line 209: first \widehat{\Sigma} should be \Sigma
% Equation 63: '\Sigma' should be '\widehat{\Sigma}
% Equation 65: Why two hats?
% Line 299: 'interest' should be 'interested'
% Page 14: 'dependant' should be 'dependent'
We have corrected the paper according to all of these comments.

% \bibliographystyle{}
% \bibliography{}

\end{document}