### 2007-07-03 paper/trunk/systemfit_statistics.tex
added some references that describe simultaneous equation systems
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@784 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Statistical background}\label{sec:statistics}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section we give a short overview of the statistical background
on that the \pkg{systemfit} package is based on.
More detailed descriptions of simultaneous equations systems
are available for instance in
\citet[chapter~7]{theil71}
\citet[part~4]{judge82}
\citet[part~5]{judge85}
\citet{srivastava87},
\citet[chapters 14--15]{greene03}, and
\citet[chapter~10]{zivot06}.
After introducing notations and assumptions,
we provide the formulas to estimate systems of linear equations.
We then demonstrate how to impose linear restrictions on parameters.
Finally, we present additional relevant issues about estimation of
equation systems.

Consider a system of $G$ equations, where the $i$th equation is of
the form 
\begin{equation}
   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G ,
\end{equation}
where $y_i$ is a vector of the dependent variable,
$X_i$ is a matrix of the exogenous variables,
$\beta_i$ is the coefficient vector and
$u_i$ is a vector of the disturbance terms of the $i$th equation.

We can write the °stacked° system as
\begin{equation}
   \left[ \begin{array}{c}
      y_1 \\ y_2\\ \vdots\\ y_G
   \end{array} \right] = 
   \left[ \begin{array}{cccc}
      X_1 & 0 & \cdots & 0\\
      0 & X_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & X_G
   \end{array}\right]
   \left[ \begin{array}{c}
      \beta_1 \\ \beta_2 \\ \vdots\\ \beta_G
   \end{array} \right] +
   \left[ \begin{array}{c}
      u_1 \\ u_2 \\ \vdots\\ u_G
   \end{array} \right]
\end{equation}

or more simply as
\begin{equation}
   y = X \beta + u .
\end{equation}

We assume that there is no correlation of the disturbance terms 
across observations, so that
\begin{equation}
   \E \left[ u_{it} \, u_{js} \right] = 0
   \; \forall \; t \neq s ,
\end{equation}
where $i$ and $j$ indicate the equation number 
and $t$ and $s$ denote the observation number.

However, we explicitly allow for contemporaneous correlation, i.e.\
\begin{equation}
   \E \left[ u_{it} \, u_{jt} \right] = \sigma_{ij} .
\end{equation}

Thus, the covariance matrix of all disturbances is
\begin{equation}
   \E \left[ u \, u' \right] = \Omega = \Sigma \otimes I ,
\end{equation}
where $\Sigma = \left[ \sigma_{ij} \right]$ is the (contemporaneous)
disturbance covariance matrix and $I$ is an identity matrix.

\input{systemfit_estimation.tex}
\input{systemfit_restrictions.tex}
\input{systemfit_other.tex}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "systemfit"
%%% End: