### 2007-12-03 paper/trunk/systemfit_compare_lm.tex
a minor improvement and a minor correction in the appendix comparing the objects returned by systemfit and lm
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@868 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\section[Object returned by systemfit]{Object returned by \code{systemfit}}
\label{sec:returned-object}

\code{systemfit} returns a list of class \code{systemfit}
that contains the results that belong to the entire system
of equations.
One special element of this list is called \code{eq}.
It is a list that contains one object for each estimated equation.
These objects are of the class \code{systemfit.equation}
and contain the results that belong only to the regarding equation.

\begin{table}[htbp]
\caption{Elements returned by \code{systemfit} and \code{lm}}
\label{tab:compare-lm}
\centering
\setlength{\tabcolsep}{4mm}
{\ttfamily
\begin{tabular}{lll}
\hline  \hline
\textbf{\code{lm}} & \textbf{\code{systemfit}} & \textbf{\code{systemfit.equation}} \\
\hline
coefficients    & coefficients   & coefficients \\
                & coefCov        & coefCov \\
fitted.values   &                & fitted.values \\
residuals       &                & residuals \\
                & residCov       & \\
                & residCovEst    & \\
rank            & rank           & rank \\
                &                & rank.sys \\
                &                & nCoef.sys \\
df.residual     & df.residual    & df.residual \\
                &                & df.residual.sys \\
call            & call           & \\
terms           &                & terms \\
                &                & inst \\
weights         &                & \\
contrasts       &                & \\
xlevels         &                & \\
offset          &                & \\
model*          &                & modelFrame* \\
y**             &                & response** \\
x**             &                & modelMatrix** \\
                &                & instMatrix** \\
                & iter           & \\
                & eq             & \\
                &                & eqnLabel \\
                &                & eqnNo \\
                & method         & method \\
                & panelLike      & \\
                & restrict.matrix& \\
                & restrict.rhs   & \\
                & restrict.regMat& \\
                & control        & \\
\hline  \hline
\end{tabular}
}
\medskip\\
Note:
\texttt{*)} if requested by the user (default is \code{TRUE})\\
\texttt{**)} if requested by the user (default is \code{FALSE})
\end{table}

The elements returned by \code{systemfit}
are similar to those returned by \code{lm},
the basic tool for linear regressions in \proglang{R}.
While some counterparts of elements returned by \code{lm}
can be found directly in objects of class \code{systemfit},
other counterparts are available for each equation
in objects of class \code{systemfit.equation}.
This is demonstrated in table~\ref{tab:compare-lm}.