### 2007-12-05 paper/trunk/systemfit_compare_lm.tex
updated paper, because arguments 'returnModelFrame', 'returnModelMatrix', 'returnResponse', and 'returnInstMatrix' of 'systemfit() have been renamed as 'model', 'x', 'y', and 'z', respectively, and elements 'modelFrame', 'modelMatrix', 'response', and 'instMatrix' returned by 'systemfit()' have been renamed as 'model', 'x', 'y', and 'z', respectively (revision 506--513 of the systemfit package); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@874 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\section[Object returned by systemfit]{Object returned by \code{systemfit}}
\label{sec:returned-object}

\code{systemfit} returns a list of class \code{systemfit}
that contains the results that belong to the entire system
of equations.
One special element of this list is called \code{eq}.
It is a list that contains one object for each estimated equation.
These objects are of the class \code{systemfit.equation}
and contain the results that belong only to the regarding equation.

\begin{table}[htbp]
\caption{Elements returned by \code{systemfit} and \code{lm}}
\label{tab:compare-lm}
\centering
\setlength{\tabcolsep}{4mm}
{\ttfamily
\begin{tabular}{lll}
\hline  \hline
\textbf{\code{lm}} & \textbf{\code{systemfit}} & \textbf{\code{systemfit.equation}} \\
\hline
coefficients    & coefficients   & coefficients \\
                & coefCov        & coefCov \\
fitted.values   &                & fitted.values \\
residuals       &                & residuals \\
                & residCov       & \\
                & residCovEst    & \\
rank            & rank           & rank \\
                &                & rank.sys \\
                &                & nCoef.sys \\
df.residual     & df.residual    & df.residual \\
                &                & df.residual.sys \\
call            & call           & \\
terms           &                & terms \\
                &                & inst \\
weights         &                & \\
contrasts       &                & \\
xlevels         &                & \\
offset          &                & \\
model*          &                & model* \\
y**             &                & y** \\
x**             &                & x** \\
                &                & z** \\
                & iter           & \\
                & eq             & \\
                &                & eqnLabel \\
                &                & eqnNo \\
                & method         & method \\
                & panelLike      & \\
                & restrict.matrix& \\
                & restrict.rhs   & \\
                & restrict.regMat& \\
                & control        & \\
\hline  \hline
\end{tabular}
}
\medskip\\
Note:
\texttt{*)} if requested by the user (default is \code{TRUE})\\
\texttt{**)} if requested by the user (default is \code{FALSE})
\end{table}

The elements returned by \code{systemfit}
are similar to those returned by \code{lm},
the basic tool for linear regressions in \proglang{R}.
While some counterparts of elements returned by \code{lm}
can be found directly in objects of class \code{systemfit},
other counterparts are available for each equation
in objects of class \code{systemfit.equation}.
This is demonstrated in table~\ref{tab:compare-lm}.