### 2007-12-29 paper/trunk/systemfit_abstract.tex
modifications of Achim Zeileis (editor of the JSS) just before publication (mainly small cosmetic improvements)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@939 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\Abstract{
Many statistical analyses (e.g., in econometrics, biostatistics and
experimental design) are based on models containing systems of
structurally related equations.
The \pkg{systemfit} package provides the capability
to estimate systems of linear equations
within the \proglang{R} programming environment.
For instance, this package can be used for 
``ordinary least squares'' (OLS),
``seemingly unrelated regression'' (SUR),
and the instrumental variable (IV) methods
``two-stage least squares'' (2SLS) and
``three-stage least squares'' (3SLS),
where SUR and 3SLS estimations can optionally be iterated.
Furthermore, the \pkg{systemfit} package provides tools
for several statistical tests.
It has been tested on a variety of datasets
and its reliability is demonstrated.
}

\Keywords{\proglang{R}, system of simultaneous equations,
  seemingly unrelated regression, two-stage least squares,
  three-stage least squares, instrumental variables}
\Plainkeywords{R, system of simultaneous equations,
  seemingly unrelated regression, two-stage least squares,
  three-stage least squares, instrumental variables}

%% at least one keyword must be supplied