### 2008-09-12 paper/trunk/systemfit.Rnw
the master file (systemfit.tex) is an Sweave file now (systemfit.Rnw) (although it not yet contains R code but the subfiles should be moved into this master file soon)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@947 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%       $Id$

\documentclass[article]{jss}
\usepackage{csquotes}
\MakeOuterQuote{°}
\usepackage{amsmath}
\usepackage{float}

% \usepackage{lineno}
% 
% \linenumbers

\newcommand{\bHat}{\hat{\beta}}
\newcommand{\COVHat}{\widehat{\COV}}
\newcommand{\lHat}{\hat{\lambda}}
\newcommand{\OHat}{\widehat{\Omega}}
\newcommand{\SHat}{\widehat{\Sigma}}
\newcommand{\sHat}{\hat{\sigma}}
\newcommand{\uHat}{\hat{u}}
\newcommand{\XHat}{\widehat{X}}
\newcommand{\tr}{\mathrm{tr}}
\newcommand{\LR}{\mathit{LR}}
\clubpenalty=10000
\widowpenalty=10000
\setlength{\emergencystretch}{3em}
% \setlength{\overfullrule}{5pt}

\newcommand{\codeD}[2]{\code{#1}\hspace{0pt}\code{.#2}}
\newcommand{\codeDD}[3]{\code{#1}\hspace{0pt}\code{.#2}\hspace{0pt}\code{.#3}}
\newcommand{\codeDDD}[4]{\code{#1}\hspace{0pt}\code{.#2}\hspace{0pt}%
\code{.#3}\hspace{0pt}\code{.#4}}

\hyphenation{systemfit}

\author{Arne Henningsen\\University of Kiel
   \And Jeff D.\ Hamann\\ Forest Informatics, Inc.}
\Plainauthor{Arne Henningsen, Jeff D. Hamann}

\title{\pkg{systemfit}: A Package for Estimating Systems of Simultaneous Equations in \proglang{R}}
\Plaintitle{systemfit: A Package for Estimating Systems of Simultaneous Equations in R}
\Shorttitle{\pkg{systemfit}: Estimating Systems of Simultaneous Equations in \proglang{R}}

%% an abstract and keywords
\input{systemfit_abstract.tex}

%% publication information
%% NOTE: This needs to filled out ONLY IF THE PAPER WAS ACCEPTED.
%% If it was not (yet) accepted, leave them commented.
\Volume{23}
\Issue{4}
\Month{December}
\Year{2007}
\Submitdate{2006-03-15}
\Acceptdate{2007-10-24}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
   Arne Henningsen\\
   Department of Agricultural Economics\\
   University of Kiel\\
   D-24098 Kiel, Germany\\
   E-mail: \email{ahenningsen@agric-econ.uni-kiel.de}\\
   URL: \url{http://www.uni-kiel.de/agrarpol/ahenningsen/}\\
   \\
   Jeff D.\ Hamann\\
   Forest Informatics, Inc.\\
   PO Box 1421\\
   Corvallis, Oregon 97339-1421, United States of America\\
   E-mail: \email{jeff.hamann@forestinformatics.com}\\
   URL: \url{http://www.forestinformatics.com/}\\
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/1/31336-5053
%% Fax: +43/1/31336-734

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\input{systemfit_intro.tex}
\input{systemfit_statistics.tex}
\input{systemfit_code.tex}
\input{systemfit_usage.tex}
\input{systemfit_reliability.tex}
\input{systemfit_summary.tex}

\section*{Acknowledgments}
We thank Achim Zeileis, John Fox, Ott Toomet, William H.\ Greene,
two anonymous referees
and several users of \pkg{systemfit}
for their comments and suggestions
that helped us to improve the \pkg{systemfit} package as well as this paper.
Of course, any remaining errors are the authors'.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{systemfit} % a subset of my big bibtex file
%\bibliography{agrarpol}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

\begin{appendix}

\input{systemfit_compare_lm.tex}

\input{systemfit_timings.tex}

\input{systemfit_sem.tex}

\end{appendix}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: