### 2008-09-12 paper/trunk/systemfit.Rnw
moved section "Using systemfit" (systemfit_usage.Snw) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@955 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%       $Id$

\documentclass[article]{jss}
\usepackage{csquotes}
\MakeOuterQuote{°}
\usepackage{amsmath}
\usepackage{float}

% \usepackage{lineno}
% 
% \linenumbers

\newcommand{\bHat}{\hat{\beta}}
\newcommand{\COVHat}{\widehat{\COV}}
\newcommand{\lHat}{\hat{\lambda}}
\newcommand{\OHat}{\widehat{\Omega}}
\newcommand{\SHat}{\widehat{\Sigma}}
\newcommand{\sHat}{\hat{\sigma}}
\newcommand{\uHat}{\hat{u}}
\newcommand{\XHat}{\widehat{X}}
\newcommand{\tr}{\mathrm{tr}}
\newcommand{\LR}{\mathit{LR}}
\clubpenalty=10000
\widowpenalty=10000
\setlength{\emergencystretch}{3em}
% \setlength{\overfullrule}{5pt}

\newcommand{\codeD}[2]{\code{#1}\hspace{0pt}\code{.#2}}
\newcommand{\codeDD}[3]{\code{#1}\hspace{0pt}\code{.#2}\hspace{0pt}\code{.#3}}
\newcommand{\codeDDD}[4]{\code{#1}\hspace{0pt}\code{.#2}\hspace{0pt}%
\code{.#3}\hspace{0pt}\code{.#4}}

\hyphenation{systemfit}

\author{Arne Henningsen\\University of Kiel
   \And Jeff D.\ Hamann\\ Forest Informatics, Inc.}
\Plainauthor{Arne Henningsen, Jeff D. Hamann}

\title{\pkg{systemfit}: A Package for Estimating Systems of Simultaneous Equations in \proglang{R}}
\Plaintitle{systemfit: A Package for Estimating Systems of Simultaneous Equations in R}
\Shorttitle{\pkg{systemfit}: Estimating Systems of Simultaneous Equations in \proglang{R}}

%% an abstract and keywords
\Abstract{
Many statistical analyses (e.g., in econometrics, biostatistics and
experimental design) are based on models containing systems of
structurally related equations.
The \pkg{systemfit} package provides the capability
to estimate systems of linear equations
within the \proglang{R} programming environment.
For instance, this package can be used for 
``ordinary least squares'' (OLS),
``seemingly unrelated regression'' (SUR),
and the instrumental variable (IV) methods
``two-stage least squares'' (2SLS) and
``three-stage least squares'' (3SLS),
where SUR and 3SLS estimations can optionally be iterated.
Furthermore, the \pkg{systemfit} package provides tools
for several statistical tests.
It has been tested on a variety of datasets
and its reliability is demonstrated.
}

\Keywords{\proglang{R}, system of simultaneous equations,
  seemingly unrelated regression, two-stage least squares,
  three-stage least squares, instrumental variables}
\Plainkeywords{R, system of simultaneous equations,
  seemingly unrelated regression, two-stage least squares,
  three-stage least squares, instrumental variables}

%% at least one keyword must be supplied

%% publication information
%% NOTE: This needs to filled out ONLY IF THE PAPER WAS ACCEPTED.
%% If it was not (yet) accepted, leave them commented.
\Volume{23}
\Issue{4}
\Month{December}
\Year{2007}
\Submitdate{2006-03-15}
\Acceptdate{2007-10-24}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
   Arne Henningsen\\
   Department of Agricultural Economics\\
   University of Kiel\\
   D-24098 Kiel, Germany\\
   E-mail: \email{ahenningsen@agric-econ.uni-kiel.de}\\
   URL: \url{http://www.uni-kiel.de/agrarpol/ahenningsen/}\\
   \\
   Jeff D.\ Hamann\\
   Forest Informatics, Inc.\\
   PO Box 1421\\
   Corvallis, Oregon 97339-1421, United States of America\\
   E-mail: \email{jeff.hamann@forestinformatics.com}\\
   URL: \url{http://www.forestinformatics.com/}\\
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/1/31336-5053
%% Fax: +43/1/31336-734

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Many theoretical models that are econometrically estimated
consist of more than one equation.
The disturbance terms of these equations are likely
to be contemporaneously correlated,
because unconsidered factors
that influence the disturbance term in one equation
probably influence the disturbance terms in other equations, too.
Ignoring this contemporaneous correlation
and estimating these equations separately
leads to inefficient estimates of the coefficients.
However, estimating all equations simultaneously
with a °generalized least squares° (GLS) estimator,
which takes the covariance structure of the residuals into account,
leads to efficient estimates.
This estimation procedure is generally called
°seemingly unrelated regression° \citep[SUR,][]{zellner62}.
Another reason to estimate a system of equations simultaneously are
cross-equation restrictions on the coefficients.%
\footnote{
Especially the economic theory suggests many cross-equation
restrictions on the coefficients (e.g., the symmetry restriction in demand models).
}
Estimating the coefficients under cross-equation restrictions
and testing these restrictions
requires a simultaneous estimation approach.

Furthermore, these models can contain variables
that appear on the left-hand side in one equation
and on the right-hand side of another equation.
Ignoring the endogeneity of these variables can lead to inconsistent
estimates.
This simultaneity bias can be corrected for by
applying a °two-stage least squares° (2SLS) estimation
to each equation.
Combining this estimation method with the SUR method results
in a simultaneous estimation of the system of equations
by the °three-stage least squares° (3SLS) method
\citep{zellner62b}.

% For all of the methods developed in the package, the disturbances of
% the individual equations are assumed to be independent and identically
% distributed (iid).  
% In the future, we would like to add the ability to fit equations were
% the disturbances are serially correlated (wikins 1969).

The \pkg{systemfit} package provides the capability to estimate
systems of linear equations in \proglang{R}
\citep{r-project}.
Currently, the estimation methods
°ordinary least squares° (OLS),
°weighted least squares° (WLS),
°seemingly unrelated regression° (SUR),
°two-stage least squares° (2SLS),
°weighted two-stage least squares° (W2SLS), and
°three-stage least squares° (3SLS)
are implemented.%
\footnote{
In this context,
the term °weighted° in °weighted least squares° (WLS) and
°weighted two-stage least squares° (W2SLS) means
that the \emph{equations} might have different weights
and \emph{not}
that the \emph{observations} have different weights.
}
The WLS, SUR, W2SLS, and 3SLS estimates can be based
either on one-step (OLS or 2SLS) (co)variances
or these estimations can be iterated,
where the (co)variances are calculated from the estimates of the previous step.
Furthermore,
the \pkg{systemfit} package
provides statistical tests for restrictions on the coefficients
and for testing the consistency of the 3SLS estimation.

Although systems of linear equations can be estimated
with several other statistical and econometric software packages
(e.g., \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
\pkg{systemfit} has several advantages.
First, all estimation procedures are publicly available in the source code.
Second, the estimation algorithms can be easily modified to meet specific
requirements.
Third, the (advanced) user can control estimation details generally
not available in other software packages by overriding reasonable defaults.

% This paper is organized as follows:
In Section~\ref{sec:statistics} we introduce the statistical
background of estimating equation systems.
The implementation of the statistical procedures in \proglang{R} is
briefly explained in Section~\ref{sec:code}.
Section~\ref{sec:Usage} demonstrates how to run
\pkg{systemfit} and how some of the features presented in the second
section can be used.
In Section~\ref{sec:reliability} we replicate several textbook results
with the \pkg{systemfit} package.
Finally, a summary and outlook are presented in
Section~\ref{sec:Summmary}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Statistical background}\label{sec:statistics}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section we give a short overview of the statistical background
that the \pkg{systemfit} package is based on.
More detailed descriptions of simultaneous equations systems
are available for instance in
\citet[Chapter~7]{theil71},
\citet[Part~4]{judge82},
\citet[Part~5]{judge85},
\citet{srivastava87},
\citet[Chapters 14--15]{greene03}, and
\citet[Chapter~10]{zivot06}.

After introducing notations and assumptions,
we provide the formulas to estimate systems of linear equations.
We then demonstrate how to estimate coefficients under linear restrictions.
Finally, we present additional relevant issues about estimation of
equation systems.

Consider a system of $G$ equations, where the $i$th equation is of
the form 
\begin{equation}
   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G ,
   \label{eq:model}
\end{equation}
where $y_i$ is a vector of the dependent variable,
$X_i$ is a matrix of the exogenous variables,
$\beta_i$ is the coefficient vector and
$u_i$ is a vector of the disturbance terms of the $i$th equation.

We can write the °stacked° system as
\begin{equation}
   \left[ \begin{array}{c}
      y_1 \\ y_2\\ \vdots\\ y_G
   \end{array} \right] = 
   \left[ \begin{array}{cccc}
      X_1 & 0 & \cdots & 0\\
      0 & X_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & X_G
   \end{array}\right]
   \left[ \begin{array}{c}
      \beta_1 \\ \beta_2 \\ \vdots\\ \beta_G
   \end{array} \right] +
   \left[ \begin{array}{c}
      u_1 \\ u_2 \\ \vdots\\ u_G
   \end{array} \right]
   \label{eq:model-array}
\end{equation}

or more simply as
\begin{equation}
   y = X \beta + u .
   \label{eq:model-matrices}
\end{equation}

We assume that there is no correlation of the disturbance terms 
across observations, so that
\begin{equation}
   \E \left[ u_{it} \, u_{js} \right] = 0
   \; \forall \; t \neq s ,
\end{equation}
where $i$ and $j$ indicate the equation number 
and $t$ and $s$ denote the observation number,
where the number of observations is the same for all equations.

However, we explicitly allow for contemporaneous correlation, i.e.,
\begin{equation}
   \E \left[ u_{it} \, u_{jt} \right] = \sigma_{ij} .
\end{equation}

Thus, the covariance matrix of all disturbances is
\begin{equation}
   \E \left[ u \, u^\top \right] = \Omega = \Sigma \otimes I_T ,
\end{equation}
where $\Sigma = \left[ \sigma_{ij} \right]$ is the (contemporaneous)
disturbance covariance matrix,
$\otimes$ is the Kronecker product,
$I_T$ is an identity matrix of dimension $T$,
and $T$ is the number of observations in each equation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimation with only exogenous regressors}
\label{sec:Estimation-ols-wls-sur}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

If all regressors are exogenous,
the system of equations (Equation~\ref{eq:model})
can be consistently estimated by ordinary least squares (OLS),
weighted least squares (WLS), and seemingly unrelated regression (SUR).
These estimators can be obtained by
\begin{equation}
   \bHat = \left( X^\top \OHat^{-1} X \right)^{-1} X^\top \OHat^{-1} y .
   \label{eq:ols-wls-sur}
\end{equation}
The covariance matrix of these estimators
can be estimated by
\begin{equation}
   \COVHat \left[ \bHat \right] = \left( X^\top \OHat^{-1} X \right)^{-1} .
   \label{eq:cov-ols-wls-sur}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Ordinary least squares (OLS)}

The ordinary least squares (OLS) estimator is based on the assumption
that the disturbance terms are not contemporaneously correlated
$(\sigma_{ij} = 0 \; \forall \; i \neq j)$
and have the same variance in each equation
$( \sigma_i^2 = \sigma_j^2 \, \forall \, i, j)$.
In this case, $\OHat$ in Equation~\ref{eq:ols-wls-sur}
is equal to $I_{G \cdot T}$ and thus, cancels out.
The OLS estimator is efficient,
as long as the disturbances are not contemporaneously correlated.

If the whole system is treated as one single equation,
$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} is $\sHat^2 I_{G \cdot T}$,
where $\sHat^2$ is an estimator for the variance of all disturbances
$(\sigma^2 = \E [ u_{it}^2 ])$.
If the disturbance terms of the individual equations
are allowed to have different variances,
$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} is $\SHat \otimes I_T$,
where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.

If the estimated coefficients are not constrained by cross-equation restrictions,
the simultaneous
OLS estimation of the system leads to the same estimated coefficients
as an equation-wise OLS estimation.
The covariance matrix of the coefficients from an equation-wise
OLS estimation is equal to the covariance matrix obtained
by Equation~\ref{eq:cov-ols-wls-sur}
with $\OHat$ equal to $\SHat \otimes I_T$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Weighted least squares (WLS)}

The weighted least squares (WLS) estimator allows for different variances
of the disturbance terms in the different equations
$( \sigma_i^2 \neq \sigma_j^2 \, \forall \, i \neq j)$,
but assumes
that the disturbance terms are not contemporaneously correlated.
In this case, $\OHat$ in Equations~\ref{eq:ols-wls-sur}
and~\ref{eq:cov-ols-wls-sur}
is $\SHat \otimes I_T$,
where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance terms in the $i$th equation.
Theoretically, $\sHat_{ii}$ should be the variance of
the (true) disturbances $( \sigma_{ii} )$.
However, they are not known in most empirical applications.
Therefore, true variances are generally replaced by estimated variances
$( \sHat_{ii} )$
that are calculated from the residuals of a first-step OLS estimation
(see Section~\ref{sec:residcov}).%
\footnote{%
Note that $\OHat$ in Equation~\ref{eq:ols-wls-sur}
is not the same $\OHat$ as in Equation~\ref{eq:cov-ols-wls-sur}.
The first is calculated from the residuals of a first-step OLS estimation;
the second is calculated from the residuals of this WLS estimation.
The same applies to the SUR, W2SLS, and 3SLS estimations
described in the following sections.
}

The WLS estimator is (asymptotically) efficient only
if the disturbance terms are not contemporaneously correlated.
If the estimated coefficients are not constrained by cross-equation restrictions,
they are equal to OLS estimates.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Seemingly unrelated regression (SUR)}

If the disturbances are contemporaneously correlated, a generalized
least squares (GLS) estimation leads to an efficient estimator for the coefficients.
In this case, the GLS estimator is generally called °seemingly unrelated regression°
(SUR) estimator \citep{zellner62}.
However, the true covariance matrix of the disturbance terms
is generally unknown.
The textbook solution for this problem is
a feasible generalized least squares (FGLS) estimation.
As the FGLS estimator is based on
an estimated covariance matrix of the disturbance terms,
it is only asymptotically efficient.
In case of a SUR estimator, $\OHat$ in Equations~\ref{eq:ols-wls-sur}
and~\ref{eq:cov-ols-wls-sur}
is $\SHat \otimes I_T$,
where $\SHat$ is the estimated covariance matrix of the disturbance terms.

It should be noted that while an unbiased OLS or WLS estimation requires only that
the regressors and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i^\top X_i \right] = 0 \; \forall \; i )$,
a consistent SUR estimation requires that all disturbance terms and all 
regressors are uncorrelated
$( \E \left[ u_i^\top X_j \right] = 0 \; \forall \; i, j )$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimation with endogenous regressors}
\label{sec:Estimation-2sls-w2sls-3sls}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

If the regressors of one or more equations are correlated
with the disturbances ($\E \left[ u_i^\top X_i \right] \neq 0$),
OLS, WLS, and SUR estimates are biased.
This can be circumvented by a
two-stage least squares (2SLS),
weighted two-stage least squares (W2SLS),
or a three-stage least squares (3SLS) estimation
with instrumental variables (IV).
The instrumental variables for each equation $Z_i$ 
can be either different or identical for all equations.
They must not be correlated with
the disturbance terms of the corresponding equation 
($\E \left[ u_i^\top Z_i \right] = 0$).

At the first stage new (°fitted°) regressors are obtained by
\begin{equation}
   \XHat_i = Z_i \left( Z_i^\top Z_i \right)^{-1} Z_i^\top X_i .
\end{equation}
Then, these °fitted° regressors are substituted for the original regressors
in Equation~\ref{eq:ols-wls-sur}
to obtain unbiased 2SLS, W2SLS, or 3SLS estimates of $\beta$ by
\begin{equation}
   \bHat
   = \left( \XHat^\top \OHat^{-1} \XHat \right)^{-1} \XHat^\top \OHat^{-1} y ,
   \label{eq:2sls-w2sls-3sls}
\end{equation}
where
\begin{equation}
   \XHat =
   \left[ \begin{array}{cccc}
      \XHat_1 & 0       & \cdots & 0\\
      0       & \XHat_2 & \cdots & 0\\
      \vdots  & \vdots  & \ddots & \vdots\\
      0       & 0       & \cdots & \XHat_G
   \end{array}\right] .
\end{equation}
An estimator of the covariance matrix of the estimated coefficients
can be obtained from Equation~\ref{eq:cov-ols-wls-sur} analogously.
Hence, we get
\begin{equation}
   \COVHat \left[ \bHat \right]
   = \left( \XHat^\top \OHat^{-1} \XHat \right)^{-1} .
   \label{eq:cov-2sls-w2sls-3sls}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Two-stage least squares (2SLS)}

The two-stage least squares (2SLS) estimator
is based on the same assumptions about the disturbance terms as the OLS estimator.
Accordingly, $\OHat$ in Equation~\ref{eq:2sls-w2sls-3sls}
is equal to $I_{G \cdot T}$ and thus, cancels out.
Like for the OLS estimator,
the whole system can be treated either as one single equation
with $\OHat$ in Equation~\ref{eq:cov-2sls-w2sls-3sls}
equal to $\sHat^2 I_{G \cdot T}$,
or the disturbance terms of the individual equations
are allowed to have different variances
with $\OHat$ in Equation~\ref{eq:cov-2sls-w2sls-3sls}
equal to $\SHat \otimes I_T$,
where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Weighted two-stage least squares (W2SLS)}

The weighted two-stage least squares (W2SLS) estimator allows
for different variances of the disturbance terms in the different equations.
Hence, $\OHat$ in Equations~\ref{eq:2sls-w2sls-3sls}
and~\ref{eq:cov-2sls-w2sls-3sls}
is $\SHat \otimes I_T$,
where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$.
If the estimated coefficients are not constrained by cross-equation restrictions,
they are equal to 2SLS estimates.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Three-stage least squares (3SLS)}

If the disturbances are contemporaneously correlated,
a feasible generalized least squares (FGLS) version of the two-stage least squares
estimation leads to consistent and asymptotically more efficient estimates.
This estimation procedure is generally called °three-stage least
squares° \citep[3SLS,][]{zellner62b}.
The standard 3SLS estimator and its covariance matrix are obtained by
Equations~\ref{eq:2sls-w2sls-3sls} and~\ref{eq:cov-2sls-w2sls-3sls}
with $\OHat$ equal to $\SHat \otimes I_T$,
where $\SHat$ is the estimated covariance matrix of the disturbance terms.

While an unbiased 2SLS or W2SLS estimation requires only that
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i^\top Z_i \right]) = 0 \; \forall \; i )$,
\cite{schmidt90} points out that the 3SLS estimator is only consistent
if all disturbance terms and all instrumental variables are uncorrelated
$( \E \left[ u_i^\top Z_j \right]) = 0 \; \forall \; i, j )$.
Since there might be occasions where this cannot be avoided,
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators.

One of these approaches based on instrumental variable estimation
(3SLS-IV) is
\begin{equation}
   \bHat_\text{3SLS-IV} = \left( \XHat^\top \OHat^{-1} X
   \right)^{-1} \XHat^\top \OHat^{-1} y .
   \label{eq:3slsIv}
\end{equation}
An estimator of the covariance matrix of the estimated 3SLS-IV coefficients is
\begin{equation}
   \COVHat \left[ \bHat_\text{3SLS-IV} \right] = \left( \XHat^\top \OHat^{-1}
   X \right)^{-1} .
\end{equation}
Another approach based on the generalized method of moments (GMM)
estimator (3SLS-GMM) is
\begin{equation}
   \bHat_\text{3SLS-GMM} = \left( X^\top Z \left( Z^\top \OHat Z \right)^{-1}
   Z^\top X \right)^{-1} X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top y
   \label{eq:3slsGmm}
\end{equation}
with
\begin{equation}
   Z =
   \left[ \begin{array}{cccc}
      Z_1 & 0 & \cdots & 0\\
      0 & Z_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & Z_G
   \end{array}\right] .
\end{equation}
An estimator of the covariance matrix of the estimated 3SLS-GMM coefficients is
\begin{equation}
   \COVHat \left[ \bHat_\text{3SLS-GMM} \right] =
   \left( X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top X \right)^{-1} .
\end{equation}
A fourth approach developed by \cite{schmidt90} himself is
\begin{equation}
   \bHat_\text{3SLS-Schmidt} = \left( \XHat^\top \OHat^{-1} \XHat
   \right)^{-1} \XHat^\top \OHat^{-1}
   Z \left( Z^\top Z \right)^{-1} Z^\top y .
   \label{eq:3slsSchmidt}
\end{equation}
An estimator of the covariance matrix of these estimated coefficients is
\begin{align}
   \COVHat \left[ \bHat_\text{3SLS-Schmidt} \right] = &
   \left( \XHat^\top \OHat^{-1}  \XHat \right)^{-1}
   \XHat^\top \OHat^{-1} Z \left( Z^\top Z \right)^{-1} Z^\top \OHat Z \\
   & \left( Z^\top Z \right)^{-1} Z^\top \OHat^{-1} \XHat
   \left( \XHat^\top \OHat^{-1}  \XHat \right)^{-1} .
   \nonumber
\end{align}
The econometrics software \proglang{EViews} uses
\begin{equation}
   \bHat_\text{3SLS-EViews} = \bHat_\text{2SLS} +
   \left( \XHat^\top \OHat^{-1} \XHat \right)^{-1}
   \XHat^\top \OHat^{-1} \left( y - X \bHat_\text{2SLS} \right) ,
   \label{eq:3slsEViews}
\end{equation}
where $\bHat_\text{2SLS}$ is the two-stage least squares estimator
as defined above.
\proglang{EViews} uses the standard 3SLS formula
(Equation~\ref{eq:cov-2sls-w2sls-3sls}) to
calculate an estimator of the covariance matrix of the estimated coefficients.

If the same instrumental variables are used in all equations 
($Z_1 = Z_2 = \ldots = Z_G$),
all the above mentioned approaches lead to identical estimates.
However, if this is not the case, the results depend on the 
method used \citep{schmidt90}.
The only reason to use different instruments for different equations
is a correlation of the instruments of one equation with the
disturbance terms of another equation.
Otherwise, one could simply use all instruments in every equation
\citep{schmidt90}.
In this case, only the 3SLS-GMM (Equation~\ref{eq:3slsGmm})
and the 3SLS estimator developed by \cite{schmidt90} 
(Equation~\ref{eq:3slsSchmidt}) are consistent.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimation under linear restrictions on the coefficients}
\label{sec:Restrictions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In many empirical applications,
it is desirable to estimate the coefficients under linear restrictions.
For instance, in econometric demand and production analysis,
it is common to estimate the coefficients
under homogeneity and symmetry restrictions
that are derived from the underlying theoretical model.

There are two different methods to estimate the coefficients
under linear restrictions.
First, a matrix $M$ can be specified that
\begin{equation}
   \beta = M \cdot \beta^\text{M} \label{eq:T-restr} ,
\end{equation}
where $\beta^\text{M}$ is a vector of restricted (linear independent) coefficients,
and $M$ is a matrix with the number of rows equal to the number of
unrestricted coefficients ($\beta$) and
the number of columns equal to the number of restricted coefficients
($\beta^\text{M}$).
$M$ can be used to map each unrestricted coefficient to one or more
restricted coefficients.

The second method to estimate the coefficients under linear restrictions
constrains the coefficients by
\begin{equation}
   R \beta^\text{R} = q ,
   \label{eq:restr-R}
\end{equation}
where $\beta^\text{R}$ is the vector of the restricted coefficients,
and $R$ and $q$ are a matrix and vector, respectively,
that specify the restrictions \citep[see][p.~100]{greene03}.
Each linear independent restriction is represented by one row of $R$
and the corresponding element of~$q$.

The first method is less flexible than the second%
\footnote{
While restrictions like $\beta_1 = 2 \beta_2$ can be specified by
both methods,
restrictions like $\beta_1 + \beta_2 = 4$ can be specified only
by the second method.
}, 
but is preferable if the coefficients are estimated
under many equality constraints across different equations of the system.
Of course, these restrictions can be also specified using
the latter method.
However, while the latter method increases the dimension of the 
matrices to be inverted during estimation, the first reduces it. 
Thus, in some cases the latter way leads to estimation problems
(e.g., (near) singularity of the matrices to be inverted),
while the first does not.

These two methods can be combined.
In this case, the restrictions specified using the latter method are
imposed on the linear independent coefficients
that are restricted by the first method, so that
\begin{equation}
   R \beta^\text{MR} = q ,
\end{equation}
where $\beta^\text{MR}$ is the vector of the restricted $\beta^\text{M}$ coefficients.

\subsubsection{Calculation of restricted estimators}

If the first method (Equation~\ref{eq:T-restr}) is chosen
to estimate the coefficients under these restrictions,
the matrix of regressors $X$ is (post-)\hspace{0pt}multiplied
by the $M$ matrix,
so that
\begin{equation}
    X^\text{M} = X \cdot M .
\end{equation}
Then, $X^\text{M}$ is substituted for $X$ and a standard estimation as described
in the previous section is done
(Equations~\ref{eq:ols-wls-sur}--\ref{eq:3slsEViews}).
This results in the linear independent coefficient estimates $\bHat^\text{M}$ and
their covariance matrix.
The original coefficients can be obtained by Equation~\ref{eq:T-restr}
and the estimated covariance matrix of the original coefficients
can be obtained by
\begin{equation}
   \COVHat \left[ \bHat \right]
   = M \cdot \COVHat \left[ \bHat^\text{M} \right] \cdot M^\top .
\end{equation}

The implementation of the second method to estimate the coefficients
under linear restrictions (Equation~\ref{eq:restr-R})
is described for each estimation method in the following sections.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Restricted OLS, WLS, and SUR estimation}

The OLS, WLS, and SUR estimators restricted by $R \beta^\text{R} = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^\text{R} \\[0.2em] \lHat
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X^\top \OHat^{-1} X & R^\top \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X^\top \OHat^{-1} y \\ q
   \end{array} \right] ,
   \label{eq:ols-wls-sur-r}
\end{equation}
where $\lambda$ is a vector of the Lagrangean multipliers of the restrictions
and $\OHat$ is defined as in Section~\ref{sec:Estimation-ols-wls-sur}.
An estimator of the covariance matrix of the estimated coefficients is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^\text{R} \\[0.2em] \lHat
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X^\top \OHat^{-1} X & R^\top \\
      R & 0
   \end{array} \right]^{-1} .
   \label{eq:cov-ols-wls-sur-r}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Restricted 2SLS, W2SLS, and 3SLS estimation}

The 2SLS, W2SLS, and standard 3SLS estimators
restricted by $R \beta^\text{R} = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^\text{R} \\[0.2em] \lHat
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \XHat^\top \OHat^{-1} \XHat & R^\top \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \XHat^\top \OHat^{-1} y \\ q
   \end{array} \right] ,
   \label{eq:2sls-w2sls-3sls-r}
\end{equation}
where $\OHat$ is defined as in Section~\ref{sec:Estimation-2sls-w2sls-3sls}.
An estimator of the covariance matrix of the estimated coefficients is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^\text{R} \\[0.2em] \lHat
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \XHat^\top \OHat^{-1} \XHat & R^\top \\
      R & 0
   \end{array} \right]^{-1} .
   \label{eq:cov-2sls-w2sls-3sls-r}
\end{equation}

The 3SLS-IV estimator restricted by $R \beta^\text{R} = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^\text{R}_\text{3SLS-IV} \\[0.2em] \lHat
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \XHat^\top \OHat^{-1} X & R^\top \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \XHat^\top \OHat^{-1} y \\ q
   \end{array} \right] ,
   \label{eq:3slsIvR}
\end{equation}
where
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^\text{R}_\text{3SLS-IV} \\[0.2em] \lHat
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \XHat^\top \OHat^{-1} X & R^\top \\
      R & 0
   \end{array} \right]^{-1} .
\end{equation}
The restricted 3SLS-GMM estimator can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^\text{R}_\text{3SLS-GMM} \\[0.2em] \lHat
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top X & R^\top \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X^\top Z \left( Z \OHat Z \right)^{-1} Z^\top y \\ q
   \end{array} \right] ,
   \label{eq:3slsGmmR}
\end{equation}
where
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^\text{R}_\text{3SLS-GMM} \\[0.2em] \lHat
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top X & R^\top \\
      R & 0
   \end{array} \right]^{-1} .
\end{equation}
The restricted 3SLS estimator based on the suggestion of
\cite{schmidt90} is
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^\text{R}_\text{3SLS-Schmidt} \\[0.2em] \lHat
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \XHat^\top \OHat^{-1} \XHat & R^\top \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \XHat^\top \OHat^{-1} Z \left( Z^\top Z \right)^{-1} Z^\top y \\ q
   \end{array} \right] ,
   \label{eq:3slsSchmidtR}
\end{equation}
where
\begin{eqnarray}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^\text{R}_\text{3SLS-Schmidt} \\[0.2em] \lHat
   \end{array} \right] 
   & = & 
   \left[ \begin{array}{cc}
      \XHat^\top \OHat^{-1} \XHat & R^\top \\
      R & 0
   \end{array} \right]^{-1}
   \\
   & & \cdot
   \left[ \begin{array}{cc}
      \XHat^\top \OHat^{-1} Z \left( Z^\top Z \right)^{-1} Z^\top \OHat
      Z \left( Z^\top Z \right)^{-1} Z^\top \OHat^{-1} \XHat & 0^\top \\
      0 & 0
   \end{array} \right]^{-1}
   \nonumber \\
   & & \cdot
   \left[ \begin{array}{cc}
      \XHat^\top \OHat^{-1} \XHat & R^\top \\
      R & 0
   \end{array} \right]^{-1} .
   \nonumber
\end{eqnarray}
The econometrics software \proglang{EViews} calculates the restricted 3SLS estimator by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^\text{R}_\text{3SLS-EViews} \\[0.2em] \lHat
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \XHat^\top \OHat^{-1} \XHat & R^\top \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \XHat^\top \OHat^{-1} \left( y - X \bHat^\text{R}_\text{2SLS} \right)
      \\ q 
   \end{array} \right] ,
   \label{eq:3slsEViewsR}
\end{equation}
where $\bHat^\text{R}_\text{2SLS}$ is the restricted 2SLS estimator calculated
by Equation~\ref{eq:2sls-w2sls-3sls-r}.
\proglang{EViews} uses the standard formula of the restricted 3SLS
estimator (Equation~\ref{eq:cov-2sls-w2sls-3sls-r}) to calculate an estimator
for the covariance matrix of the estimated coefficients.

If the same instrumental variables are used in all equations 
($Z_1 = Z_2 = \ldots = Z_G$),
all the above mentioned approaches lead to identical coefficient estimates
and identical covariance matrices of the estimated coefficients.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Other issues and tools}\label{sec:Other}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Residual covariance matrix}\label{sec:residcov}

Since the (true) disturbances ($u_i$) of the estimated equations
are generally not known,
their covariance matrix cannot be determined.
Therefore, this covariance matrix is generally calculated from
estimated residuals ($\uHat_i$)
that are obtained from a first-step OLS or 2SLS estimation.
Then, in a second step, the estimated residual covariance matrix
can be employed for a WLS, SUR, W2SLS, or 3SLS estimation.
In many cases, the residual covariance matrix is calculated by
\begin{equation}
   \sHat_{ij} = \frac{ \uHat_i^\top \uHat_j }{ T },
   \label{eq:rcovNoDfCor}
\end{equation}
where $T$ is the number of observations in each equation.
However, in finite samples this estimator is biased,
because it is not corrected for degrees of freedom.
The usual single-equation procedure to correct for degrees of freedom
cannot always be applied, because the number of regressors in each equation
might differ.
Two alternative approaches to calculate the residual covariance
matrix are
\begin{equation}
   \sHat_{ij} = \frac{ \uHat_i^\top \uHat_j }
   { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
   \label{eq:rcovGeomean}
\end{equation}
and
\begin{equation}
   \sHat_{ij} = \frac{ \uHat_i^\top \uHat_j }
   { T - \max \left( K_i , K_j \right) } \; ,
   \label{eq:rcovMax}
\end{equation}
where $K_i$ and $K_j$ are the number of regressors in equation
$i$ and $j$, respectively.
However, these formulas yield unbiased estimators only if $K_i = K_j$
\citep[p.~469]{judge85}.
% Greene (2003, p. 344) says that the second is unbiased if i = j or K_i = K_j,
% whereas the first is unbiased only if i = j. 
% However, if K_i = K_j the first and the second are equal.
% Why is the first biased if K_i = K_j ???????????

A further approach to obtain a residual covariance
matrix is
\begin{eqnarray}
   \sHat_{ij} & = &
   \frac{ \uHat_i^\top \uHat_j }
   { T - K_i - K_j + \tr \left[ X_i \left( X_i^\top X_i \right)^{-1}
   X_i^\top X_j \left( X_j^\top X_j \right)^{-1} X_j^\top \right] }
   \label{eq:rcovTheil} \\
   & = &
   \frac{ \uHat_i^\top \uHat_j }
   { T - K_i - K_j + \tr \left[ \left( X_i^\top X_i \right)^{-1}
   X_i^\top X_j \left( X_j^\top X_j \right)^{-1} X_j^\top X_i \right] }
\end{eqnarray} 
\citep[p.~309]{zellner62c}.
This yields an unbiased estimator for all elements of $\Sigma$,
but even if $\SHat$ is an unbiased estimator of $\Sigma$,
its inverse $\SHat^{-1}$ is not an unbiased estimator
of $\Sigma^{-1}$ \citep[p.~322]{theil71}.
Furthermore, the covariance matrix calculated by Equation~\ref{eq:rcovTheil}
is not necessarily positive semidefinite \citep[p.~322]{theil71}.
Hence, °it is doubtful whether [this formula] is really superior to 
[Equation~\ref{eq:rcovNoDfCor}]° \citep[p.~322]{theil71}.

The WLS, SUR, W2SLS and 3SLS coefficient estimates are consistent
if the residual covariance matrix is calculated
using the residuals from a first-step OLS or 2SLS estimation.
There exists also an alternative slightly different approach
that consists of three steps.%
\footnote{
For instance, this approach is applied by
the command \code{TSCS} of the software \proglang{LIMDEP} that carries out SUR estimations
in which all coefficient vectors are constrained to be equal
\citep{greene06}.
}
In a first step, an OLS or 2SLS estimation is applied to obtain residuals
to calculate a (first-step) residual covariance matrix.
In a second step, the first-step residual covariance matrix
is used to estimate the model by WLS or W2SLS
and new residuals are obtained to calculate
a (second-step) residual covariance matrix.
Finally, in the third step,
the second-step residual covariance matrix is used
to estimate the model by SUR or 3SLS.
If the estimated coefficients are not constrained by cross-equation restrictions,
OLS and WLS estimates as well as 2SLS and W2SLS estimates are identical.
Hence, in this case both approaches generate the same results.

It is also possible to iterate WLS, SUR, W2SLS and 3SLS estimations.
At each iteration the residual covariance matrix is calculated
from the residuals of the previous iteration.
If Equation~\ref{eq:rcovNoDfCor} is applied to calculate the
residual covariance matrix,
an iterated SUR estimation converges to maximum
likelihood \citep[p.~345]{greene03}.

In some uncommon cases,
for instance in pooled estimations,
where the coefficients are restricted to be equal in all equations,
the means of the residuals of each equation are not equal to zero
$( \overline{ \uHat }_i \neq 0 )$.
Therefore, it might be argued
that the residual covariance matrix should be calculated
by subtracting the means from the residuals
and substituting $\uHat_i - \overline{ \uHat }_i$
for $\uHat_i$ in Equations~\ref{eq:rcovNoDfCor}--\ref{eq:rcovTheil}.

If the coefficients are estimated under any restrictions,
the residual covariance matrix
for a WLS, SUR, W2SLS, or 3SLS estimation
can be obtained either from a restricted or from an unrestricted
first-step estimation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Degrees of freedom}
\label{sec:degreesOfFreedom}

To our knowledge the question about how to determine the degrees
of freedom for single-coefficient $t$ tests is not comprehensively
discussed in the literature.
While sometimes the degrees of freedom of the entire system
(total number of observations in all equations minus
total number of estimated coefficients)
are applied,
in other cases the degrees of freedom of each single equation
(number of observations in the equations minus
number of estimated coefficients in the equation)
are used.
Asymptotically, this distinction does not make a difference.
However, in many empirical applications, the number of observations
of each equation is rather small, and
therefore, it matters.

If a system of equations is estimated by an unrestricted OLS and
the covariance matrix of the coefficients is calculated
with $\OHat$ in Equation~\ref{eq:cov-ols-wls-sur}
equal to $\SHat \otimes I_T$,
the estimated coefficients and their standard errors are identical
to an equation-wise OLS estimation.
In this case, it is reasonable to use the degrees of freedom of
each single equation,
because this yields the same $P$ values as the equation-wise
OLS estimation.

In contrast, if a system of equations is estimated with many
cross-equation restrictions and
the covariance matrix of an OLS estimation is calculated
with $\OHat$ in Equation~\ref{eq:cov-ols-wls-sur}
equal to $\sHat^2 I_{G \cdot T}$,
the system estimation is similar to a single equation estimation.
Therefore, in this case, it seems to be reasonable to use the degrees
of freedom of the entire system.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Goodness of fit}

The goodness of fit of each single equation can be measured by the
traditional $R^2$ values
\begin{equation}
   R_i^2 = 1 - \frac{ \uHat_i^\top \uHat_i }
   { ( y_i - \overline{y_i} )^\top ( y_i - \overline{y_i} ) } \; ,
\end{equation}
where $R_i^2$ is the $R^2$ value of the $i$th equation
and $\overline{y_i}$ is the mean value of $y_i$.

The goodness of fit of the whole system can be measured by the
McElroy's $R^2$ value 
% also: \citep[p.~345]{greene03}
\begin{equation}
   R_*^2 = 1 - \frac{ \uHat^\top \OHat^{-1} \uHat }
   { y^\top \left( \SHat^{-1} \otimes
   \left( I_T - \frac{\iota \iota^\top}{T} \right) \right) y } ,
\end{equation}
where $\iota$ is a column vector of $T$ ones \citep{mcelroy77}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing linear restrictions}
\label{sec:testingRestrictions}

Linear restrictions can be tested by an $F$ test, two Wald tests and
a likelihood ratio (LR) test.

The $F$ statistic for systems of equations is
\begin{equation}
F = \frac{
   ( R \bHat - q )^\top
   ( R ( X^\top ( \Sigma \otimes I )^{-1} X )^{-1} R^\top )^{-1}
   ( R \bHat - q ) /
   j
}{
   \uHat^\top ( \Sigma \otimes I )^{-1} \uHat /
   ( G \cdot T - K )
} \; ,
\label{eq:f-test}
\end{equation}
where $j$ is the number of restrictions,
$K$ is the total number of estimated coefficients,
and all other variables are as defined before
\citep[p.~314]{theil71}.
Under the null hypothesis, $F$ is $F$ distributed
with $j$ and $G \cdot T - K$ degrees of freedom.

However, $F$ in Equation~\ref{eq:f-test} cannot be computed,
because $\Sigma$ is generally unknown.
As a solution, \citet[p.~314]{theil71} proposes
to replace the unknown $\Sigma$ in Equation~\ref{eq:f-test}
by the estimated covariance matrix $\SHat$.
\begin{equation}
\hat{F} = \frac{
   ( R \bHat - q )^\top
   ( R ( X^\top ( \SHat \otimes I )^{-1} X )^{-1} R^\top )^{-1}
   ( R \bHat - q ) /
   j
}{
   \uHat^\top ( \SHat \otimes I )^{-1} \uHat /
   ( G \cdot T - K )
}
\label{eq:f-test-theil}
\end{equation}
Asymptotically, $\hat{F}$ has the same distribution as $F$
in Equation~\ref{eq:f-test},
because the numerator of Equation~\ref{eq:f-test-theil}
converges in probability to the numerator of Equation~\ref{eq:f-test}
and the denominator of Equation~\ref{eq:f-test-theil}
converges in probability to the denominator of Equation~\ref{eq:f-test}
\citep[p.~402]{theil71}.

Furthermore, the denominators of both Equations~\ref{eq:f-test}
and~\ref{eq:f-test-theil} converge in probability to~$1$.
Taking this into account and applying Equation~\ref{eq:cov-ols-wls-sur},
we obtain the usual $F$~statistic of the Wald test.
\begin{equation}
\hat{\hat{F}} = \frac{
   ( R \bHat - q )^\top
   ( R \, \COVHat \left[ \bHat \right] R^\top )^{-1}
   ( R \bHat - q )
}{
   j
}
\label{eq:f-test-wald}
\end{equation}
Under the null hypotheses, also $\hat{\hat{F}}$ is
asymptotically $F$ distributed
with $j$ and $G \cdot T - K$ degrees of freedom.

Multiplying Equation~\ref{eq:f-test-wald} with $j$,
we obtain the usual $\chi^2$ statistic for the Wald test
\begin{equation}
W =
   ( R \bHat - q )^\top
   ( R \, \COVHat [ \bHat ] \, R^\top )^{-1}
   ( R \bHat - q ) .
\label{eq:chi2-test-wald}
\end{equation}
Asymptotically, $W$ has a $\chi^2$
distribution with $j$ degrees of freedom
under the null hypothesis
\citep[p.~347]{greene03}.

The likelihood-ratio (LR) statistic for systems of equations is
\begin{equation}
\LR = T \cdot \left(
   \log \left| \SHat_r \right|
   - \log \left| \SHat_u \right|
   \right) ,
\label{eq:lr-test}
\end{equation}
where $T$ is the number of observations per equation, and 
$\SHat_r$ and $\SHat_u$ are
the residual covariance matrices calculated by Equation~\ref{eq:rcovNoDfCor}
of the restricted and unrestricted estimation, respectively.
Asymptotically, $\LR$ has a $\chi^2$
distribution with $j$ degrees of freedom
under the null hypothesis
\citep[p.~349]{greene03}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hausman test}
\label{sec:hausman}

\citet{hausman78} developed a test for misspecification.
The null hypothesis of the test is
that the instrumental variables of each equation are
uncorrelated with the disturbance terms of all other equations
($\E \left[ u_i^\top Z_j \right] = 0 \, \forall \, i \neq j$).
Under this null hypothesis, both the 2SLS and the 3SLS estimator are consistent,
but the 3SLS estimator is (asymptotically) more efficient.
Under the alternative hypothesis, the 2SLS estimator is consistent
but the 3SLS estimator is inconsistent,
i.e., the instrumental variables of each equation are
uncorrelated with the disturbances of the same equation
($\E \left[ u_i^\top Z_i \right] = 0 \, \forall \, i$),
but the instrumental variables of at least one equation
are correlated with the disturbances of another equation
($\E \left[ u_i^\top Z_j \right] \neq 0 \, \exists \, i \neq j$).
The Hausman test statistic is
\begin{equation}
  m = \left( \bHat_\text{2SLS} - \bHat_\text{3SLS} \right)^\top
      \left( \COVHat \left[ \bHat_\text{2SLS} \right] -
             \COVHat \left[ \bHat_\text{3SLS} \right] \right)
      \left( \bHat_\text{2SLS} - \bHat_\text{3SLS} \right) ,
\label{eq:hausman}
\end{equation}
where $\bHat_\text{2SLS}$ and $\COVHat \left[ \bHat_\text{2SLS} \right]$ are the estimated
coefficient and covariance matrix from a 2SLS estimation, and
$\bHat_\text{3SLS}$ and $\COVHat \left[ \bHat_\text{3SLS} \right]$ are the estimated
coefficients and covariance matrix from a 3SLS estimation.
Under the null hypothesis, this test statistic has a
$\chi^2$ distribution with degrees of freedom equal to the number of
estimated coefficients.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Source code}\label{sec:code}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<echo=FALSE>>=
options( prompt = "R> " )
@

The source code of the \pkg{systemfit} package is publicly available
for download from the Comprehensive \proglang{R} Archive Network (CRAN,
\url{http://CRAN.R-project.org/}).
The basic functionality of this package is provided by the
function \code{systemfit}.
Moreover, this package provides
tools for statistical tests,
functions (methods) to show, extract or calculate results,
some convenience functions,
and internal helper functions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection[The basic function systemfit]{The basic function \code{systemfit}}

The \code{systemfit} function estimates systems of linear equations
by different estimation methods.
Where possible,
the user interface and the returned object of this function
follow the function \code{lm}
--- the basic tool for linear regressions in \proglang{R} ---
to make the usage of \code{systemfit} as easy as possible
for \proglang{R} users.

The econometric estimation is done by applying
the formulas in Sections~\ref{sec:Estimation-ols-wls-sur}
and~\ref{sec:Estimation-2sls-w2sls-3sls}
or --- if the coefficients are estimate under linear restrictions ---
by the formulas in Section~\ref{sec:Restrictions}.
If the restrictions on the coefficients are specified symbolically,
function \code{makeHypothesis} of the \pkg{car} package \citep{r-car-1.2-1, fox02a}
is used to create the restriction matrix.

The \code{systemfit} function returns a list of class \code{systemfit}
that contains the results that belong to the entire system of equations.
One special element of this list is called \code{eq},
which is a list that contains one object for each estimated equation.
These objects are lists of class \codeD{systemfit}{equation}
and contain the results that belong only to the regarding equation.
A complete description is available in the documentation of this function
that is included in the package.
A comparison of the elements returned by \code{lm} and by \code{systemfit}
is available in appendix~\ref{sec:returned-object}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Statistical tests}

The \codeD{linear}{hypothesis} and
\code{lrtest} methods for \code{systemfit} objects
as well as the function \codeD{hausman}{systemfit}
apply the statistical tests described in
Sections~\ref{sec:testingRestrictions} and~\ref{sec:hausman}.

The \codeD{linear}{hypothesis} method for {systemfit} objects
can be used to test linear restrictions on the estimated coefficients
by \citeauthor{theil71}'s $F$ test or by usual Wald tests.
Internally, \citeauthor{theil71}'s $F$ statistic is computed
by the hidden function \codeD{.ftest}{systemfit}
and the Wald tests are computed by
the default \codeD{linear}{hypothesis} method of the \pkg{car} package
\citep{r-car-1.2-1, fox02a}.
The \code{lrtest} method for \code{systemfit} objects is a wrapper function
to the default \code{lrtest} method of the \pkg{lmtest} package
\citep{r-lmtest},
which computes the likelihood-ratio (LR) test statistic.
All these functions return an object of class \code{anova}
that contains --- amongst others ---
the empirical test statistic, the degrees of freedom,
and the corresponding $P$ value.

The function \codeD{hausman}{systemfit}
tests the consistency of the 3SLS estimator.
It returns an object of class \code{htest},
which contains --- amongst others ---
the empirical test statistic,
the degrees of freedom, and the $P$ value.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Other methods and functions}

The \pkg{systemfit} package provides several methods
for objects both of classes \code{systemfit}
and \codeD{systemfit}{equation}:
\code{print} methods print the estimation results,
\code{summary} methods calculate summary results,
\code{confint} methods compute confidence intervals for the coefficients,
\code{predict} methods calculate predicted values,
\code{coef} methods extract the estimated coefficients,
\code{vcov} methods extract their covariance matrix,
\code{fitted} methods extract the fitted values,
\code{residuals} methods extract the residuals,
\code{formula} methods extract the formula(s),
\code{terms} methods extract the model terms,
\codeD{model}{frame} methods extract the model frame,
and \codeD{model}{matrix} methods extract the model matrix.
Some methods can be applied to objects of class \code{systemfit} only:
a \code{correlation} method calculates
the correlations between the predictions of two equations,
an \code{se.ratio} method computes the ratios of the standard errors
of the predictions between two models,
and a \code{logLik} method extracts the log likelihood value.
The package provides \code{print} methods
to print objects of classes
\codeD{summary}{systemfit}, \codeDD{summary}{systemfit}{equation},
and \codeD{confint}{systemfit}
that are returned by the above mentioned \code{summary}
and \code{confint} methods.
There exist also two \code{coef} methods
to extract the estimated coefficients,
their standard errors, $t$ values, and $P$ values
from objects of classes
\codeD{summary}{systemfit} and \codeDD{summary}{systemfit}{equation}.%
\footnote{%
There does not exist a special method to extract
the degrees of freedom of the residuals from \code{systemfit} objects,
because the default method of \code{df.residual}
works for these objects.
}

The convenience function
\code{createSystemfitModel} creates a model for \code{systemfit}
by random numbers;
\codeD{systemfit}{control} sets suitable default values
for the technical control parameters for \code{systemfit}.

Finally, the package includes some internal (hidden) helper functions:
\codeD{.prepareData}{systemfit},
\code{.stackMatList}, and \code{.prepareWmatrix}
for preparing the data matrices;
\code{.calcXtOmegaInv} and \code{.calcGLS}
for calculating the GLS estimator;
\code{.calcResidCov} and \code{.calcSigma2}
for calculating the (co)variances of the residuals;
and \codeD{.ftest}{systemfit}
for calculating \citeauthor{theil71}'s $F$ statistic.
If \code{systemfit} is applied to
a (classical) °seemingly unrelated regression° analysis with panel data,
it calls the hidden internal function \code{.systemfitPanel},
which reshapes the data, creates the formulas to be estimated,
and --- if requested --- specifies restrictions
to ensure that the coefficients of all individuals are equal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiency of computations}
\label{sec:code-efficiency}
We have followed \cite{bates04} to make the code of \pkg{systemfit}
faster and more stable.
First, if a formula contains an inverse of a matrix
that is post-multiplied by a vector or matrix,
we use \code{solve(A,b)} instead of \code{solve(A) \%*\% b}.
Second, we calculate crossproducts
by \code{crossprod(X)} or \code{crossprod(X,y)}
instead of \code{t(X) \%*\% X} or \code{t(X) \%*\% y},
respectively.

The matrix $\Omega^{-1}$ that is used to compute the estimated
coefficients and their covariance matrix is of size
$( G \cdot T ) \times ( G \cdot T )$
(see Sections~\ref{sec:Estimation-ols-wls-sur},
\ref{sec:Estimation-2sls-w2sls-3sls}, and~\ref{sec:Restrictions}).
In case of large data sets, $\Omega^{-1}$ becomes computationally infeasible.
Therefore, we use the following transformation and compute $X^\top \Omega^{-1}$
by dividing the $X$ matrix into submatrices,
doing some calculations with these submatrices,
adding up some of these submatrices, and
finally putting the submatrices together, so that
\begin{equation}
X^\top \Omega^{-1}
%= X^\top \left( \Sigma^{-1} \otimes I \right)
= \sum_{i=1} \left[ \begin{array}{c}
   \sigma^{1i} {X^1} \\
   \sigma^{2i} {X^2} \\
   \vdots \\
   \sigma^{Gi} {X^G} \\
\end{array} \right]^\top ,
\label{eq:omegaInv}
\end{equation}
where $\sigma^{ij}$ are the elements of the matrix $\Sigma^{-1}$,
and $X^i$ is a submatrix of $X$ that contains the rows
that belong to the $i$'s equation.
This computation is done inside the internal (hidden) function
\code{.calcXtOmegaInv}.

Since version 1.0, the \code{systemfit} function by default
uses the \pkg{Matrix} package \citep{r-matrix-07}
for all computations where matrices are involved.
The \pkg{Matrix} package provides classes for different types of matrices.
For instance, we choose class \code{dgeMatrix}
(°real matrices in general storage mode°),
for matrices $X_i$ in Equation~\ref{eq:model-array},
class \code{dgCMatrix} (°general, numeric, sparse matrices
in the (sorted) compressed sparse column format°)
for matrix $X$ in Equation~\ref{eq:model-matrices},
and class \code{dspMatrix}
(°symmetric real matrices in packed storage (one triangle only)°)
for the residual covariance matrix $\SHat$.
If the \pkg{Matrix} package is used,
the possibly huge matrix $\Omega^{-1}$ is no longer a problem,
because it is a sparse matrix
that can be stored in a compressed format (class \code{dgCMatrix}).
Hence, we no longer need the algorithm in Equation~\ref{eq:omegaInv}.
We have tested different ways to calculate a GLS estimator
like in Equation~\ref{eq:ols-wls-sur}
and we found
that the following code is the fastest:
<<eval=FALSE>>=
sigmaInv <- solve( residCov )
xtOmegaInv <- crossprod( xMat, kronecker( sigmaInv, Diagonal( nObs ) ) )
coef <- solve( xtOmegaInv %*% xMat, xtOmegaInv %*%  yVec )
@
In this code snippet,
\code{residCov} is the residual covariance matrix $\SHat$,
\code{nObs} is the number of observations in each equation $T$,
\code{xMat} is the matrix $X$ and
\code{yVec} is the vector $y$ in Equation~\ref{eq:ols-wls-sur}.

By default, the \code{systemfit} function uses the \pkg{Matrix} package
to perform GLS estimations,
because using this package considerably decreases
the computation time for many common models.
However, the estimation of small models with small data sets
gets slower by using the \pkg{Matrix} package
(see appendix~\ref{sec:timings}).
While this increase in computation time
is often imperceptible to human beings,
it might matter in some cases
such as iterated estimations or Monte Carlo studies.
Therefore, the user can opt for not using the \pkg{Matrix} package,
but Equation~\ref{eq:omegaInv} with standard \proglang{R} matrices.

\subsection[Overlap with other functions and packages in R]
   {Overlap with other functions and packages in \proglang{R}}

Single-equation models can be fitted in \proglang{R}
by OLS with function \code{lm} (package \pkg{stats})
and by 2SLS with function \code{tsls}
(package \pkg{sem}, \citealp{r-sem-0.9-8}).
This is also possible with the \code{systemfit} function,
but \code{systemfit} is specialized in estimating systems of equation,
i.e., more than one equation.
Its capability to estimate single-equation models is just a side-effect.

Function \code{sem} (package \pkg{sem}, \citealp{r-sem-0.9-8})
can be used to estimate structural equation models in \proglang{R}
by limited information maximum likelihood (LIML)
and full information maximum likelihood (FIML)
assuming normal or multinormal errors, respectively.
A special feature of this function is the estimation
of models with unobserved (°latent°) variables,
which is not possible with \code{systemfit}.
While \code{sem} cannot be used to consistently estimate
systems of simultaneous equations with some endogenous regressors,
it can be used to estimate systems of equations,
where all regressors are exogenous.
However, the latter is rather cumbersome
(see appendix~\ref{sec:sem}).
Hence, \code{systemfit} is the only function in \proglang{R}
that can be used to estimate systems of simultaneous equations
and it is the most convenient function to estimate
systems of equations with purely exogenous regressors.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Using systemfit]{Using \pkg{systemfit}}\label{sec:Usage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SweaveOpts{prefix.string=trash/usage}
<<echo=FALSE>>=
options( prompt = "R> " )
@

In this section we demonstrate how to use the \pkg{systemfit} package.
First, we show the standard usage of \code{systemfit} by a simple example.
Second, several options that can be specified by the user are presented.
Then, the usage of \code{systemfit}
for a (classical) °seemingly unrelated regression° analysis
with panel data is described.
Finally, we demonstrate how to apply some statistical tests.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection[Standard usage of systemfit]{Standard usage of \code{systemfit}}
\label{sec:standard-usage}

As described in the previous section,
systems of equations can be econometrically estimated
with the function \code{systemfit}.
The only mandatory argument is \code{formula}.
Typically, it is a list of formulas to be estimated,
but it may also be a single formula for estimating a single-equation model.
Each formula is a standard regression formula in \proglang{R}
(see documentation of \code{formula}).

The following demonstration uses an example taken from
\citet[p.~685]{kmenta86}.
We want to estimate a small model of the US food market:
\begin{align}
\texttt{consump} & = \beta_1 + \beta_2 \cdot \texttt{price}
   + \beta_3 \cdot \texttt{income} \\
\texttt{consump} & = \beta_4 + \beta_5 \cdot \texttt{price}
   + \beta_6 \cdot \texttt{farmPrice} + \beta_7 \cdot \texttt{trend}
\end{align}
The first equation represents the demand side of the food market.
Variable \code{consump} (food consumption per capita) is the dependent variable.
The regressors are \code{price} (ratio of food prices to general consumer prices)
and \code{income} (disposable income) as well as a constant.
The second equation specifies the supply side of the food market.
Variable \code{consump} is the dependent variable of this equation as well.
The regressors are again \code{price} (ratio of food prices to general 
consumer prices) and a constant as well as 
\code{farmPrice} (ratio of preceding year's prices received by farmers
to general consumer prices) and 
\code{trend} (a time trend in years).
These equations can be estimated by OLS in \proglang{R} by
<<systemfit_usage>>=
library( "systemfit" )
data( "Kmenta" )
attach( Kmenta )
eqDemand <- consump ~ price + income
eqSupply <- consump ~ price + farmPrice + trend
eqSystem <- list( demand = eqDemand, supply = eqSupply )
fitols <- systemfit( eqSystem )
print( fitols )
@

The first line loads the \pkg{systemfit} package.
The second line loads example data that are included with the package.
They are attached to the \proglang{R} search path in line three.
In the fourth and fifth line, the demand and supply equations are specified,
respectively.%
\footnote{
A regression constant is always implied if not explicitly omitted.
}
In the sixth line, these equations are concatenated into a list and
are labeled \code{demand} and \code{supply}, respectively.%
\footnote{
If no labels are provided, the equations are numbered consecutively
( \code{eq1}, \code{eq2}, \ldots ).
}
Finally, in the last two lines, the regression is performed
and the estimation results are printed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection[User options of systemfit]{User options of \code{systemfit}}
\label{sec:user-options}

The user can modify the default estimation method by providing
additional optional arguments,
e.g., to specify instrumental variables
or restrictions on the coefficients.
All optional arguments are described in the following:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Estimation method}

The optional argument \code{method} is a string
that determines the estimation method.
It must be either \code{"OLS"}, \code{"WLS"}, \code{"SUR"},
\code{"2SLS"}, \code{"W2SLS"}, or \code{"3SLS"}.
These methods correspond to the estimation methods
described in Sections~\ref{sec:Estimation-ols-wls-sur},
\ref{sec:Estimation-2sls-w2sls-3sls}, and~\ref{sec:Restrictions}.
The following command estimates the model described above as
°seemingly unrelated regression°.
<<systemfit_usage>>=
fitsur <- systemfit( eqSystem, method = "SUR" )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Instrumental variables}
The instruments for a 2SLS, W2SLS or 3SLS estimation can be specified
by the argument \code{inst}.
If the same instruments should be used for all equations,
\code{inst} must be a one-sided formula.%
\footnote{
A one-sided formula is a standard formula in \proglang{R}
without a dependent variable.}
If different instruments should be used for each equation,
\code{inst} must be a list that contains a one-sided formula
for each equation.
The following example uses instrumental variables
to estimate the model described above
by °three-stage least squares° (3SLS).
While the first command specifies the same instruments for all equations,
the second uses different instruments:
<<systemfit_usage>>=
fit3sls  <- systemfit( eqSystem, method = "3SLS",
   inst = ~ income + farmPrice + trend )
fit3sls2 <- systemfit( eqSystem, method = "3SLS",
   inst = list( ~ farmPrice + trend, ~ income + farmPrice + trend ) )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Data}
Having all data in the global environment or attached to the search path
is often inconvenient.
Therefore, \code{systemfit} has the argument \code{data}
to specify a data frame
that contains the variables of the model.
In the following example, we use this argument to specify
that the data for the estimation
should be taken from the data frame \code{Kmenta}.
Hence, we no longer need to attach this data frame before calling \code{systemfit}:
<<systemfit_usage>>=
fitsur <- systemfit( eqSystem, method = "SUR", data = Kmenta )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Restrictions on the coefficients}
As outlined in Section~\ref{sec:Restrictions}, restrictions on the coefficients
can be specified in two ways.
One way is to use the matrix $R$ and the vector $q$
(see Section~\ref{sec:Restrictions}).
These restrictions can be specified symbolically
by argument \codeD{restrict}{matrix}
as in the generic function
\codeD{linear}{hypothesis} of the \pkg{car} package \citep{r-car-1.2-1, fox02a}.
This argument must be a vector of character strings,
where each element represents one linear restriction,
and each element must be either a linear combination of coefficients,
or a linear equation in the coefficients
(see documentation of function \codeD{linear}{hypothesis}
in the \pkg{car} package, \citealp{r-car-1.2-1, fox02a}).
We illustrate this by estimating the model under
the restriction $\beta_2 + \beta_6 = 0$.
Since the name of $\beta_2$
(coefficient of variable \code{price} in equation \code{demand})
is \code{demand\_price}
and the name of $\beta_6$
(coefficient of variable \code{farmPrice} in equation \code{supply})
is \code{supply\_farmPrice},
this restriction can be specified by
<<systemfit_usage,keep.source=TRUE>>=
restrict <- "demand_price + supply_farmPrice = 0"
fitsurRmat <- systemfit(eqSystem, method = "SUR",
   restrict.matrix = restrict)
@
\label{code:Rmat}

Alternatively, the restrictions via matrix $R$ and vector $q$
can be specified numerically.
The matrix $R$ can be specified with argument \codeD{restrict}{matrix}
and the vector $q$ with argument \codeD{restrict}{rhs}.
<<systemfit_usage,keep.source=TRUE>>=
Rmat <- matrix(0, nrow = 1, ncol = 7)
Rmat[1, 2] <- 1
Rmat[1, 6] <- 1
qvec <- c(0)
fitsurRmatNum <- systemfit(eqSystem, method = "SUR",
   restrict.matrix = Rmat, restrict.rhs = qvec)
@
The first line creates a $1 \times 7$ matrix of zeros,
where 1 is the number of restrictions and
7 is the number of unrestricted coefficients.
The following two lines specify this matrix in a way
that the multiplication with the coefficient vector
results in $ \beta_2 + \beta_6 $.
The fourth line creates a vector with a single element
that contains the right hand side of the restriction, i.e., zero.
Finally the coefficients are estimated under the restriction
$\beta_2 + \beta_6 = 0$.

The other way to specify restrictions on the coefficients is to modify the regressor matrix
by post-multiplying it with a matrix, say $M$
(see Section~\ref{sec:Restrictions}).
This kind of restriction can be specified
by setting argument \codeD{restrict}{regMat} equal to the matrix $M$.
We convert the restriction specified above to $\beta_2 = - \beta_6$,
and set $\beta_1 = \beta^\text{M}_1$, \ldots, $\beta_5 = \beta^\text{M}_5$,
$\beta_6 = - \beta^\text{M}_2$, and $\beta_7 =  \beta^\text{M}_6$.
We can do this in \proglang{R} by
<<systemfit_usage,keep.source=TRUE>>=
modRegMat <- matrix(0, nrow = 7, ncol = 6)
modRegMat[1:5, 1:5] <- diag(5)
modRegMat[6, 2] <- -1
modRegMat[7, 6] <- 1
fitsurRegMat <- systemfit(eqSystem, method = "SUR",
   restrict.regMat = modRegMat)
@
The first line creates a $7 \times 6$ matrix of zeros,
where 7 is the number of unrestricted coefficients and
6 is the number of restricted coefficients.
The following three lines specify the matrix $M$ (\code{modRegMat})
as described before.
Finally the coefficients are estimated under the restriction
$\beta^\text{M}_2 = \beta_2 = - \beta_6$.

Of course, the estimation results do not depend on the method
that was used to specify this restriction:
<<systemfit_usage>>=
all.equal( coef( fitsurRmat ), coef( fitsurRmatNum ) )
all.equal( coef( fitsurRmat ), coef( fitsurRegMat ) )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Iteration control}
The estimation methods WLS, SUR, W2SLS and 3SLS need a
covariance matrix of the residuals
that can be calculated from a first-step OLS or 2SLS estimation
(see Section~\ref{sec:residcov}).
This procedure can be iterated and at each iteration the covariance
matrix is calculated from the previous step estimation.
This iteration is repeated until the maximum number of iterations
is reached or the coefficient estimates have converged.
The maximum number of iterations is specified by argument \code{maxiter}.
Its default value is one, which means no iteration.
The convergence criterion is
\begin{equation}
   \sqrt{ \frac{ \sum_i (\beta_{i,g} - \beta_{i,g-1})^2 }
   { \sum_i \beta_{i,g-1}^2 }}
      < \texttt{tol} ,
\end{equation}
where $\beta_{i,g}$ is the $i$th coefficient of the $g$th iteration.
The default value of the convergence criterion (argument \code{tol})
is $10^{-5}$.

In the following example,
we estimate the model described above by iterated SUR:
<<systemfit_usage>>=
fitsurit <- systemfit( eqSystem, method = "SUR", maxiter = 500 )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Residual covariance matrix}
It was explained in Section~\ref{sec:residcov} that several different
methods have been proposed to calculate the residual
covariance matrix.
The user can specify, which method \code{systemfit} should use.
Possible values of the argument \code{methodResidCov} are presented in
Table~\ref{tab:methodResidCov}.
By default, \code{systemfit} uses Equation~\ref{eq:rcovGeomean}.

\begin{table}[H]
\centering
\begin{tabular}{|c|c|}
\hline
argument & equation \\
\code{methodResidCov} & \\
\hline
\code{"noDfCor"} & \ref{eq:rcovNoDfCor} \\
\hline
\code{"geomean"} & \ref{eq:rcovGeomean} \\
\hline
\code{"max"}     & \ref{eq:rcovMax} \\
\hline
\code{"Theil"}   & \ref{eq:rcovTheil} \\
\hline
\end{tabular}
\caption{Possible values of argument \code{methodResidCov}}
\label{tab:methodResidCov}
\end{table}

Furthermore, the user can specify
whether the means should be subtracted from the residuals
before Equations~\ref{eq:rcovNoDfCor}, \ref{eq:rcovGeomean}, \ref{eq:rcovMax},
or~\ref{eq:rcovTheil} are applied
to calculate the residual covariance matrix
(see Section~\ref{sec:residcov}).
The corresponding argument is called \code{centerResiduals}.
It must be either \code{TRUE} (subtract the means) or
\code{FALSE} (take the unmodified residuals).
The default value of \code{centerResiduals} is \code{FALSE}.

Moreover, if the coefficients are estimated under restrictions,
the user can use argument \code{residCovRestricted} to specify
whether the residual covariance matrix
for a WLS, SUR, W2SLS, or 3SLS estimation
should be obtained from a restricted or from an unrestricted
first-step estimation
(see Section~\ref{sec:residcov}).
If this argument is \code{TRUE} (the default),
the residual covariance matrix is obtained from a restricted
OLS or 2SLS estimation.
If it is \code{FALSE},
the residual covariance matrix is obtained from an unrestricted
first-step estimation.

Finally, argument \code{residCovWeighted} can be used to decide,
whether the residual covariance matrix for a SUR (3SLS) estimation
should be obtained from a WLS (W2SLS) estimation
instead of from an OLS (2SLS) estimation
(see Section~\ref{sec:residcov}).
By default, \code{residCovWeighted} is \code{FALSE},
which means that the residuals of an OLS (2SLS) estimation are used
to compute the residual covariance matrix.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{3SLS formula}
As discussed in Sections~\ref{sec:Estimation-2sls-w2sls-3sls} and~\ref{sec:Restrictions},
there exist several different
methods to perform a 3SLS estimation.
The user can specify the method by argument \code{method3sls}.
Possible values are presented in Table~\ref{tab:method3sls}.
The default value is \code{"GLS"}.

\begin{table}[H]
\centering
\begin{tabular}{|c|c|c|}
\hline
argument           & equation       & equation \\
\code{method3sls} & (unrestricted) & (restricted) \\
\hline
\code{"GLS"}     & \ref{eq:2sls-w2sls-3sls} & \ref{eq:2sls-w2sls-3sls-r} \\
\hline
\code{"IV"}      & \ref{eq:3slsIv}      & \ref{eq:3slsIvR} \\
\hline
\code{"GMM"}     & \ref{eq:3slsGmm}     & \ref{eq:3slsGmmR} \\
\hline
\code{"Schmidt"} & \ref{eq:3slsSchmidt} & \ref{eq:3slsSchmidtR} \\
\hline
\code{"EViews"}  & \ref{eq:3slsEViews}  & \ref{eq:3slsEViewsR} \\
\hline
\end{tabular}
\caption{Possible values of argument \code{method3sls}}
\label{tab:method3sls}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Sigma squared}
In case of OLS or 2SLS estimations, argument \code{singleEqSigma}
can be used to specify,
whether different $\sigma^2$s for each single equation
or the same $\sigma^2$ for all equations should be used.
If argument \code{singleEqSigma} is \code{TRUE},
$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} or~\ref{eq:cov-2sls-w2sls-3sls}
is set to $\SHat \otimes I_T$.
In contrast, if argument \code{singleEqSigma} is \code{FALSE},
$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} or~\ref{eq:cov-2sls-w2sls-3sls}
is set to $\sHat^2 I_{G \cdot T}$.
In case of an unrestricted regression,
argument \code{singleEqSigma} is \code{TRUE} by default.
However, if the coefficients are estimated under restrictions,
this argument is \code{FALSE} by default.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{System options}
Furthermore, two options regarding some internal calculations are available.
First, argument \code{solvetol} specifies the tolerance level
for detecting linear dependencies when inverting a matrix or
calculating a determinant (using functions \code{solve} and \code{det}).
The default value depends on the used computer system and is equal to the
default tolerance level of \code{solve} and \code{det}.

Second, argument \code{useMatrix} specifies
whether the \pkg{Matrix} package \citep{r-matrix-07}
should be used for all computations where matrices are involved
(see Section~\ref{sec:code-efficiency}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Returned data objects}

Finally, the user can decide
whether \code{systemfit} should return some data objects.
Argument \code{model} indicates
whether a data frame with the data of the model
should be returned.
Its default value is \code{TRUE},
i.e., the model frame is returned.
Arguments \code{x}, \code{y},
and \code{z} specify
whether the model matrices ($X_i$),
the responses ($y_i$), and the matrices of instrumental variables ($Z_i$),
respectively, should be returned.
These three arguments are \code{FALSE} by default,
i.e., these data objects are not returned.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection[Summary results with summary.systemfit]
   {Summary results with \code{summary.systemfit}}

The \code{summary} method can be used
to compute and print summary results of objects
returned by \code{systemfit}.
<<systemfit_usage>>=
summary( fitsur )
@
First, the estimation method is reported
and a few summary statistics for the entire system
and for each equation are given.
Then, the covariance matrix used for estimation
and the covariance matrix as well as the correlation matrix
of the (final) residuals are printed.
Finally, the estimation results of each equation are reported:
the formula of the estimated equation,
the estimated coefficients, their standard errors, $t$ values, $P$ values
and codes indicating their statistical significance,
as well as some other statistics like
the standard error of the residuals and
the $R^2$ value of the equation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection[Degrees of freedom for t tests]
   {Degrees of freedom for $t$ tests}

The \code{summary} method for \code{systemfit} objects has an optional argument \code{useDfSys}.
It selects the approach
that is applied by \code{systemfit}
to determine the degrees of freedom of $t$ tests of the estimated coefficients
(Section~\ref{sec:degreesOfFreedom}).
If argument \code{useDfSys} is \code{TRUE},
the degrees of freedom of the whole system are taken.
In contrast, if \code{useDfSys} is \code{FALSE},
the degrees of freedom of the single equation are taken.
If the coefficients are estimated under restrictions,
argument \code{useDfSys} is \code{TRUE} by default.
However, if no restrictions on the coefficient are specified,
the default value of \code{useDfSys} is \code{FALSE}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Reduce amount of printed output}

The optional arguments \code{residCov} and \code{equations}
can be used reduce the amount of the printed output.
Argument \code{residCov} specifies
whether the covariance matrix and
the correlation matrix
of the residuals are printed.
Argument \code{equations} specifies
whether summary results of each equation are printed.
By default, both arguments are \code{TRUE}.
The following command returns a sparse summary output:
<<systemfit_usage>>=
summary( fitsur, residCov = FALSE, equations = FALSE )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Panel data}

The \code{systemfit} function can also be used
for a (classical) °seemingly unrelated regression° analysis
with panel data.
For this type of analysis,
the data must be provided in a transformed data frame of class \codeD{plm}{dim}%
\footnote{
Generally, panel data can be either in °long format°
(different individuals are arranged below each other)
or in °wide format°
(different individuals are arranged next to each other).
For this analysis, the data must be in °long format°.
}
which can be created with the function \codeD{plm}{data}
from the \proglang{R} package \pkg{plm}
\citep{r-plm-0.3-1}.
In contrast to the previously described usage of \code{systemfit},
argument \code{formula} must be a single equation (object of class \code{formula}).
This formula is estimated for all individuals.

We demonstrate the application of \code{systemfit}
to panel data
using an example taken from \citet[p.~340]{greene03}
that is based on \citet{grunfeld58}.
We want to estimate a model for gross investment of 5 US firms
in the years 1935--1954:
\begin{equation}
\texttt{invest}_{it} = \beta_1 + \beta_2 \cdot \texttt{value}_{it} +
   \beta_3 \cdot \texttt{capital}_{it}
\end{equation}
where \code{invest} is the gross investment of firm $i$ in year $t$,
\code{value} is the market value of the firm at the end of the
previous year, and
\code{capital} is the capital stock of the firm at the end of the
previous year.

This model can be estimated by
<<systemfit_usage>>=
data( "GrunfeldGreene" )
library( "plm" )
GGPanel <- plm.data( GrunfeldGreene, c( "firm", "year" ) )
greeneSur <- systemfit( invest ~ value + capital, method = "SUR",
   data = GGPanel )
@
The first line loads the example data set \code{GrunfeldGreene}
that is included in the \pkg{systemfit} package.
The second line loads the \pkg{plm} package
and the following line specifies a data frame of class \codeD{plm}{dim},
where the variables \code{firm} and \code{year} indicate
the individual (cross-section) and time identifier, respectively.
Finally, a seemingly unrelated regression is performed.

The optional argument \code{pooled} is a logical variable indicating
whether the coefficients are restricted to be equal for all
individuals.
By default, this argument is set to \code{FALSE}.
The following command does a seemingly unrelated regression
of the same model as before,
but with coefficients restricted to be equal for all individuals.
<<systemfit_usage>>=
greeneSurPooled <- systemfit( invest ~ value + capital, method = "SUR",
   data = GGPanel, pooled = TRUE )
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing linear restrictions}

As described in Section~\ref{sec:testingRestrictions},
linear restrictions can be tested by an $F$ test, two Wald tests and
an LR test.
The \pkg{systemfit} package provides
the method \codeD{linear}{hypothesis} for the $F$ and Wald tests
as well as the method \code{lrtest} for LR tests.

We will now test the restriction $\beta_2 = -\beta_6$
that was specified by the matrix \code{Rmat} and the vector \code{qvec}
in the example above (p.~\pageref{code:Rmat}).
<<systemfit_usage>>=
linear.hypothesis( fitsur, Rmat, qvec, test = "FT" )

linear.hypothesis( fitsur, Rmat, qvec, test = "F" )

linear.hypothesis( fitsur, Rmat, qvec, test = "Chisq" )

lrtest( fitsurRmat, fitsur )
@
The linear restrictions are tested by \citeauthor{theil71}'s $F$ test
(Equation~\ref{eq:f-test-theil}) first,
second by the $F$ statistic of a Wald test
(Equation~\ref{eq:f-test-wald}),
third by the $\chi^2$ statistic of a Wald test
(Equation~\ref{eq:chi2-test-wald}),
and finally by an LR test
(Equation~\ref{eq:lr-test}).

The first argument of the \codeD{linear}{hypothesis} method for \code{systemfit} objects
must be an unrestricted regression
returned by \code{systemfit}.
The second and third arguments are the restriction matrix $R$ and
the optional vector $q$, as described in Section~\ref{sec:Restrictions}.
Analogously to the argument \codeD{restrict}{matrix}
of the \code{systemfit} function,
the restrictions can be specified either in matrix form or symbolically.
The optional argument \code{test} must be
a character string, \code{"FT"}, \code{"F"}, or \code{"Chisq"},
specifying whether to compute
\citeauthor{theil71}'s finite-sample $F$ test
(with approximate $F$ distribution)
the finite-sample Wald test (with approximate $F$ distribution)
or the large-sample Wald test (with asymptotic $\chi^2$ distribution).

All arguments of the \code{lrtest} method for \code{systemfit} objects
must be fitted model objects returned by \code{systemfit}.
It consecutively compares all provided fitted model objects.

All tests print a short description of the test
and the tested model objects first.
Then, a small table is printed,
where each row belongs to one (unrestricted or restricted) model.
The second row reports (amongst others)
the degree(s) of freedom of the test,
the empirical test statistic,
and the marginal level of significance ($P$ value).
Although all tests check the same hypothesis,
there is some variation of the $P$ values.
However, all tests suggest the same decision:
The null hypothesis $\beta_2 = -\beta_6$ cannot be rejected at any
reasonable level of significance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hausman test}
A Hausman test, which is described in Section~\ref{sec:hausman},
can be carried out with following commands:
<<systemfit_usage>>=
fit2sls  <- systemfit( eqSystem, method = "2SLS",
   inst = ~ income + farmPrice + trend, data = Kmenta )
fit3sls <- systemfit( eqSystem, method = "3SLS",
   inst = ~ income + farmPrice + trend, data = Kmenta )
hausman.systemfit( fit2sls, fit3sls )
@
<<echo=FALSE>>=
hausmantest <- hausman.systemfit( fit2sls, fit3sls )
@
First of all, the model is estimated by 2SLS and then by 3SLS.
Finally, in the last line
the test is carried out by the command \codeD{hausman}{systemfit}.
This function requires two arguments:
the result of a 2SLS estimation and
the result of a 3SLS estimation.
The Hausman test statistic is
\Sexpr{round( hausmantest$statistic, digits = 3)},
which has a $\chi^2$ distribution
with \Sexpr{hausmantest$parameter} degrees of freedom
under the null hypothesis.
The corresponding $P$ value is
\Sexpr{round( hausmantest$p.value, digits = 3 )}.
This shows that the null hypothesis is not rejected
at any reasonable level of significance.
Hence, we can assume that the 3SLS estimator is consistent.

\input{systemfit_reliability.tex}
\input{systemfit_summary.tex}

\section*{Acknowledgments}
We thank Achim Zeileis, John Fox, Ott Toomet, William H.\ Greene,
two anonymous referees
and several users of \pkg{systemfit}
for their comments and suggestions
that helped us to improve the \pkg{systemfit} package as well as this paper.
Of course, any remaining errors are the authors'.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{systemfit} % a subset of my big bibtex file
%\bibliography{agrarpol}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

\begin{appendix}

\input{systemfit_compare_lm.tex}

\input{systemfit_timings.tex}

\input{systemfit_sem.tex}

\end{appendix}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: